---
title: "Data Clean"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(dplyr)
library(haven)
library(tidyr)
library(ggplot2)
library(scales)
library(ggpubr)
library(ggpattern)
library(stringr)
library(here)

```

# 2. Import the raw data 

```{r rawdata, echo=FALSE}

# PROCTCAE data
pro <- read_sas(here("data","SAS Data", "pro_responses.sas7bdat"))
length(unique(pro$PTID))
# Patient clinical and demographic data
pt_clinical <- read_sas(here("data","SAS Data","pt_clinical.sas7bdat"))
length(unique(pt_clinical$PTID))
```

# 3.1. Data pre-processing

## 3.1.1 Endpoints (reduction,death,er,hospitalization and discontinuation) 
get the time of each visit of each endpoint; No time for treatment discontinuation
         
```{r endpoints, echo=FALSE}

# Endpoint analysis (time to event)
endpoint <- pt_clinical %>%
  mutate(C12_dose_reduction_day1  =  as.numeric(C12_dose_reduction1_date - consent_dt),
         C12_dose_reduction_day2  =  as.numeric(C12_dose_reduction2_date - consent_dt),
         C12_dose_reduction_day3  =  as.numeric(C12_dose_reduction3_date - consent_dt),
         C12_dose_reduction_day4  =  as.numeric(C12_dose_reduction4_date - consent_dt),
         C12_dose_reduction_day5  =  as.numeric(C12_dose_reduction5_date - consent_dt),
         C12_dose_reduction_day6  =  as.numeric(C12_dose_reduction6_date - consent_dt),
         C12_dose_reduction_day7  =  as.numeric(C12_dose_reduction7_date - consent_dt),
         C12_dose_reduction_day8  =  as.numeric(C12_dose_reduction8_date - consent_dt),
         C12_dose_reduction_day9  =  as.numeric(C12_dose_reduction9_date - consent_dt),
         C12_dose_reduction_day10  =  as.numeric(C12_dose_reduction10_date - consent_dt)) %>%
  mutate(C9_death_date  =  as.numeric(C9_death_date - consent_dt)) %>%
  mutate(C12_er_day1  =  as.numeric(C12_er_visit1_date - consent_dt),
         C12_er_day2  =  as.numeric(C12_er_visit2_date - consent_dt),
         C12_er_day3  =  as.numeric(C12_er_visit3_date - consent_dt),
         C12_er_day4  =  as.numeric(C12_er_visit4_date - consent_dt),
         C12_er_day5  =  as.numeric(C12_er_visit5_date - consent_dt),
         C12_er_day6  =  as.numeric(C12_er_visit6_date - consent_dt),
         C12_er_day7  =  as.numeric(C12_er_visit7_date - consent_dt),
         C12_er_day8  =  as.numeric(C12_er_visit8_date - consent_dt),
         C12_er_day9  =  as.numeric(C12_er_visit9_date - consent_dt),
         C12_er_day10  =  as.numeric(C12_er_visit10_date - consent_dt),
         C12_er_day11  =  as.numeric(C12_er_visit11_date - consent_dt),
         C12_er_day12  =  as.numeric(C12_er_visit12_date - consent_dt),
         C12_er_day13  =  as.numeric(C12_er_visit13_date - consent_dt),
         C12_er_day14  =  as.numeric(C12_er_visit14_date - consent_dt),
         C12_er_day15  =  as.numeric(C12_er_visit15_date - consent_dt),
         C12_er_day16  =  as.numeric(C12_er_visit16_date - consent_dt),
         C12_er_day17  =  as.numeric(C12_er_visit17_date - consent_dt),
         C12_er_day18  =  as.numeric(C12_er_visit18_date - consent_dt),
         C12_er_day19  =  as.numeric(C12_er_visit19_date - consent_dt),
         C12_er_day20  =  as.numeric(C12_er_visit20_date - consent_dt)) %>%
  mutate(C12_hospital_day1  =  as.numeric(C12_hospital_admission1_date - consent_dt),
         C12_hospital_day2  =  as.numeric(C12_hospital_admission2_date - consent_dt),
         C12_hospital_day3  =  as.numeric(C12_hospital_admission3_date - consent_dt),
         C12_hospital_day4  =  as.numeric(C12_hospital_admission4_date - consent_dt),
         C12_hospital_day5  =  as.numeric(C12_hospital_admission5_date - consent_dt),
         C12_hospital_day6  =  as.numeric(C12_hospital_admission6_date - consent_dt),
         C12_hospital_day7  =  as.numeric(C12_hospital_admission7_date - consent_dt),
         C12_hospital_day8  =  as.numeric(C12_hospital_admission8_date - consent_dt),
         C12_hospital_day9  =  as.numeric(C12_hospital_admission9_date - consent_dt),
         C12_hospital_day10  =  as.numeric(C12_hospital_admission10_date - consent_dt)) %>%
  select(PTID,C12_dose_reduction,starts_with("C12_dose_reduction_day"),starts_with("C12_er_day"),
         starts_with("C12_hospital_day"),C9_death,C9_death_date,C12_discontinuing_overall,C12_er,C12_hospital) 

# export the data frame for separate analysis
# write.csv(endpoint,file = "endpoint.csv",row.names = FALSE)

```

## 3.1.2 Demographics
extract the patient demographics from the clinical data set

```{r demographic, include=FALSE, echo=FALSE}

test <- pt_clinical %>%
      select(C7_ECOG_v2,C7_ECOG)
  
demo <- pt_clinical %>%
  # obtain all IV and Oral treatment information of each patient (in total 6 treatment for IV and PO)
  mutate(C7_intravenous_all_1 =  do.call(paste0, across(c("C7_intravenous_1", starts_with("C7_intravenous_1_v")))),
         C7_intravenous_all_2 =  do.call(paste0, across(c("C7_intravenous_2", starts_with("C7_intravenous_2_v")))),
         C7_intravenous_all_3 =  do.call(paste0, across(c("C7_intravenous_3", starts_with("C7_intravenous_3_v")))),
         C7_intravenous_all_4 =  do.call(paste0, across(c("C7_intravenous_4", starts_with("C7_intravenous_4_v")))),
         C7_intravenous_all_5 =  do.call(paste0, across(c("C7_intravenous_5", starts_with("C7_intravenous_5_v")))),
         C7_intravenous_all_6 =  do.call(paste0, across(c("C7_intravenous_6", starts_with("C7_intravenous_6_v")))),
         C7_oral_all_1 =  do.call(paste0, across(c("C7_oral_1", starts_with("C7_oral_1_v")))),
         C7_oral_all_2 =  do.call(paste0, across(c("C7_oral_2", starts_with("C7_oral_2_v")))),
         C7_oral_all_3 =  do.call(paste0, across(c("C7_oral_3", starts_with("C7_oral_3_v")))),
         C7_oral_all_4 =  do.call(paste0, across(c("C7_oral_4", starts_with("C7_oral_4_v")))),
         C7_oral_all_5 =  do.call(paste0, across(c("C7_oral_5", starts_with("C7_oral_5_v")))),
         C7_oral_all_6 =  do.call(paste0, across(c("C7_oral_6", starts_with("C7_oral_6_v"))))) %>%
  # ECOG was recorded in either C7_ECOG_V2 or C7_ECOG
  mutate(C7_ECOG_all = across(c("C7_ECOG_v2", "C7_ECOG"), ~ ifelse(is.na(.x), "", .x)) %>%
  as.data.frame() %>%
  do.call(paste0, .)) %>% 
  mutate(C9_death_date  =  as.numeric(C9_death_date - consent_dt)) %>%
  select(PTID, consent_dt,Age,Gender,Ethnicity,race_overall,C7_height_overall, 
         cancer,C12_dose_reduction,C7_intravenous_treatment, C7_oral_treatment,
         starts_with("C7_intravenous_all_"),starts_with("C7_oral_all_"),
         C7_ECOG_all,C9_death,C9_death_date,C12_discontinuing_overall,C12_er,C12_hospital) 

```


# 4. Data combination and imputation
combine the demographics with PRO reported patients and prelim imputation process
FormStatusDay: Days between activity start date and activity form status change date

## 4.1 data combination
```{r combination, include=FALSE, echo=FALSE}

# extract the symptom score from the pro dataset
pro2 <- pro %>% 
      select(PTID, FormStatusDay, Activity, Depression,appetite_overall, starts_with("PROCTCAE_"))
# combine the PROs with demographics
dat2 <- merge(demo, pro2, by='PTID')
```

## 4.2 data imputation
```{r, include=FALSE, echo=FALSE}
#--------------------------------------
# 1. treatment combination
#--------------------------------------
# IV treatment
dat2 <- dat2 %>%
  rowwise() %>%
  mutate(intravenous_merged = 
         str_squish(str_c(na.omit(c_across(starts_with("C7_intravenous_all_"))), collapse = " ")),
         intravenous_merged = if_else(intravenous_merged == "" | is.na(intravenous_merged),
         "None",intravenous_merged)
  ) %>%
  select(-matches("^C7_intravenous_all_")) %>%
  ungroup()

# Oral treatment
dat2 <- dat2 %>%
  rowwise() %>%
  mutate(oral_merged = 
          str_squish(str_c(na.omit(c_across(starts_with("C7_oral_all_"))), collapse = " ")),
         oral_merged = if_else(oral_merged == "" | is.na(oral_merged),
                              "None",oral_merged)
  ) %>%
  select(-matches("^C7_oral_all_")) %>%
  ungroup()

#--------------------------------------
# 2. Check duplicates columns of PROs (remove the duplicate rows--keep the rows with less missing values)
#--------------------------------------
check_col <- c("Activity", "Depression", "PROCTCAE_Pain_f", "PROCTCAE_Pain_s", "PROCTCAE_Pain_i",
                "PROCTCAE_Nausea_f", "PROCTCAE_Nausea_s", "PROCTCAE_Vomit_f", "PROCTCAE_Diarr_f",
                "PROCTCAE_Constip_s", "PROCTCAE_Dyspnea_s", "PROCTCAE_Dyspnea_i", "PROCTCAE_Insomnia_s")

dat2_clean <- dat2 %>%
        rowwise() %>%
        mutate(count = sum(is.na(c_across(all_of(check_col))))) %>% 
        group_by(PTID, FormStatusDay) %>%
        arrange(count) %>%
        slice(1) %>%
        select(-count) %>%
        ungroup()

# check duplicate again - no duplicate found
any(duplicated(dat2_clean)) 

#--------------------------------------
# 3. Classify treatment into subgroups (IV + PO)
#--------------------------------------
dat3 <- dat2_clean %>%
    mutate(
    REGIMEN_IV = case_when(
# Chemo + Immuno + Targeted
      str_detect(intravenous_merged,regex("carboplatin|etoposide|pemetrexed|docetaxel|paclitaxel|cisplatin|gemcitabine|topotecan|5FU|oxaliplatin|cyclophosphamide|doxorubicin|irinotecan|fluorouracil|cabazitaxel|trabectedin|bendamustine|eribulin|vinorelbine|temsirolimus|decitabine|fludarabine", ignore_case = TRUE)) &
        
      str_detect(intravenous_merged, regex("pembrolizumab|nivolumab|atezolizumab|ipilimumab", ignore_case = TRUE)) &
        
      str_detect(intravenous_merged,regex("bevacizumab|ramucirumab|trastuzumab|cetuximab|pertuzumab|daratumumab|panitumumab|olaratumab|elotuzumab|rituximab|blinatumamab|obinutuzumab", ignore_case = TRUE)) ~ "three combine",
      
# Chemo + Immunotherapy
      str_detect(intravenous_merged,regex("carboplatin|etoposide|pemetrexed|docetaxel|paclitaxel|cisplatin|gemcitabine|topotecan|5FU|oxaliplatin|cyclophosphamide|doxorubicin|irinotecan|fluorouracil|cabazitaxel|trabectedin|bendamustine|eribulin|vinorelbine|temsirolimus|decitabine|fludarabine", ignore_case = TRUE)) &
      
      str_detect(intravenous_merged, regex("pembrolizumab|nivolumab|atezolizumab|ipilimumab", 
      ignore_case = TRUE)) ~"Chemo_Immunotherapy",
      
# Chemo + Targeted
      str_detect(intravenous_merged,regex("carboplatin|etoposide|pemetrexed|docetaxel|paclitaxel|cisplatin|gemcitabine|topotecan|5FU|oxaliplatin|cyclophosphamide|doxorubicin|irinotecan|fluorouracil|cabazitaxel|trabectedin|bendamustine|eribulin|vinorelbine|temsirolimus|decitabine|fludarabine", ignore_case = TRUE)) &
      
      str_detect(intravenous_merged,regex("bevacizumab|ramucirumab|trastuzumab|cetuximab|pertuzumab|daratumumab|panitumumab|olaratumab|elotuzumab|rituximab|blinatumamab|obinutuzumab", ignore_case = TRUE)) ~ "Chemo_Targeted",
      
# Immuno + Targeted
      str_detect(intravenous_merged, regex("pembrolizumab|nivolumab|atezolizumab|ipilimumab", ignore_case = TRUE)) &
      str_detect(intravenous_merged,regex("bevacizumab|ramucirumab|trastuzumab|cetuximab|pertuzumab|daratumumab|panitumumab|olaratumab|elotuzumab|rituximab|blinatumamab|obinutuzumab", ignore_case = TRUE)) ~ "Immuno_Targeted",
      
# Immunotherapy only
      str_detect(intravenous_merged, regex("pembrolizumab|nivolumab|atezolizumab|ipilimumab", ignore_case = TRUE))~ "Immunotherapy only",
      
# Chemotherapy only
      str_detect(intravenous_merged,regex("carboplatin|etoposide|pemetrexed|docetaxel|paclitaxel|cisplatin|gemcitabine|topotecan|5FU|oxaliplatin|cyclophosphamide|doxorubicin|irinotecan|fluorouracil|cabazitaxel|trabectedin|bendamustine|eribulin|vinorelbine|temsirolimus|decitabine|fludarabine", ignore_case = TRUE)) ~ "Chemotherapy only",
      
#Targeted only
      str_detect(intravenous_merged,regex("bevacizumab|ramucirumab|trastuzumab|cetuximab|pertuzumab|daratumumab|panitumumab|olaratumab|elotuzumab|rituximab|blinatumamab|obinutuzumab", ignore_case = TRUE)) ~ "Targeted only",
      
      # None or unclassified
      TRUE ~ "None or Unclassified"
    )
  )

# Classify Oral treatment
dat3 <- dat3 %>%
mutate(REGIMEN_PO = case_when(
# TKI
str_detect(oral_merged, regex("ceritinib|crizotinib|alectinib|afatinib|osimertinib|erlotinib|sunitinib|sorafenib|pazopanib|axitinib|lenvatinib|caboxantinib|regorafenib|imatinib", ignore_case = TRUE))  ~ "TKI Treatment",
  
# None or unclassified
      TRUE ~ "None or Unclassified"
    )
  )

```

# 5. Convert the data set to population model format

```{r, include=FALSE, echo=FALSE}
final_df <- dat3 %>%
    # remove the treatment binary variables and merged treatment variables
    select(-oral_merged,-intravenous_merged,-C7_intravenous_treatment,-C7_oral_treatment) %>%
    # change the full PROCTCAE names to numbers (start with 004 since three other symptoms will be coded 001-003)
    rename_with(~ paste0("PROCTCAE_", sprintf("%03d", seq(from = 4, length.out = length(.)))), 
                .cols = starts_with("PROCTCAE_")) %>%  
    rename(race = race_overall, height = C7_height_overall, cancertype = cancer,
             ecog = C7_ECOG_all,time = FormStatusDay,death = C9_death,death.date = C9_death_date,
             discontinue = C12_discontinuing_overall, er = C12_er, hospital = C12_hospital,
             PROCTCAE_001 = Activity, PROCTCAE_002 = Depression,PROCTCAE_003 = appetite_overall) %>%
    rename_with(toupper) %>%
    group_by(PTID) %>%
    mutate(FIRSTTIME = first(TIME)) %>%
    ungroup() %>%
    mutate(TAFR = TIME - FIRSTTIME) %>%
    pivot_longer(
      cols = starts_with("PROCTCAE_"),
      names_to = "PROCTCAE",
      values_to = "DV"
    ) %>%
    mutate(MDV = as.numeric(is.na(DV))) %>%
    group_by(PTID,TIME) %>%
    mutate(VAR = row_number()) %>%
    ungroup() %>%
    # rearrange the columns
    select(PTID,CONSENT_DT,PROCTCAE,VAR,TIME,TAFR,DV,MDV,everything())

#write.csv(final_df,"fulldata_IRT.csv",row.names = FALSE)

```
