---
title: "IRT vs. PRO Total comparison"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(here)
library(dplyr)
library(dunn.test)
library(ggplot2)
library(ggpubr)

```

# 2. Import the raw data and transformation
Individual parameter from IRT or PRO total model
# 2.1 IRT data
```{r, echo=FALSE}
train_irt_prm <- read.csv(here("data","irt_data","ML_data","model_output","train.txt")) %>%
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 
```

# 2.2 PRO total data
```{r, echo=FALSE}
train_pro_prm <- read.csv(here("data","prototal_data","prototal_final_individual.txt")) %>%
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)
```

# 3. Statistical analysis

The aim of this analysis is to compare the parameter distribution of each cancer type and previous treatment
1. Kruskal–Wallis test to assess overall group differences.
2. If p < 0.05, proceed with pairwise post-hoc Dunn's comparisons. Adjustment for multiple testing using Holm-Bonferroni.

# 3.1 Import the demographics
```{r, echo=FALSE}

train_demo <- read.csv(here("data","IRT_data","ML_data","full_IRT_training.csv")) %>%
  group_by(PTID) %>%
  distinct(PTID,.keep_all = TRUE) %>%
  select(PTID,CANCERTYPE,IV,PO) %>%
# regroup some covariates (same as previous multivariable regression of each endpoint)
  # rename IV treatment
  mutate(IV = factor(case_when(
    IV == 1 ~ "Chemo",
    IV == 2 ~ "Target",
    IV == 3 ~ "Immuno",
    IV == 4 ~ "Combination",
    IV == 5 ~ "None or Unclassified"),
    levels = c(
      "Chemo",
      "Target",
      "Immuno",
      "Combination",
      "None or Unclassified"
    ))) %>%
  # rename PO treatment
  mutate(PO = factor(case_when(
    PO == 1 ~ "TKI",
    PO == 2 ~ "None or Unclassified"))) %>%
  # rename cancer type
  mutate(CANCERTYPE = factor(case_when(
    CANCERTYPE == 1 ~ "Thoracic",
    CANCERTYPE == 2 ~ "Breast",
    CANCERTYPE %in% c(3,9,10) ~ "Gastrointestinal", #(Colorectal/Anal,Gastro-esophageal,Small bowel and 'Pancreas, Hepatobiliary)
    CANCERTYPE %in% c(4,8) ~ "Genitourinary",
    CANCERTYPE == 5 ~ "Gynecologic",
    CANCERTYPE %in% c(6,7,99) ~ "Other tumors"),
    levels = c(
      "Thoracic",
      "Breast",
      "Gastrointestinal",
      "Genitourinary",
      "Gynecologic",
      "Other tumors"
    )))
```

# 3.2 IRT model comparsion
## 3.2.1 Cancer type

```{r}
train <- train_demo %>%
        left_join(train_irt_prm,by = "PTID")

vars <- c("BASEGI","EMAXGI","KDGI","SLPGI",
          "BASEPS","EMAXPS","KDPS","SLPPS",
          "BASEPD","EMAXPD","KDPD","SLPPD")

# Kruskal-Wallis results
krus_cancer_irt <- data.frame()
for (var in vars) {
  kw <- kruskal.test(as.formula(paste(var, "~ CANCERTYPE")), data = train)
  krus_cancer_irt <- rbind(krus_cancer_irt, data.frame(
    parameter = var,
    H = round(kw$statistic, 3),
    p_value = round(kw$p.value, 4)
  ))
}

print(krus_cancer_irt)
#parameter      H p_value
#Kruskal-Wallis chi-squared	BASEGI	15.284	0.0092	
#Kruskal-Wallis chi-squared8	BASEPD	13.484	0.0192	

# Dunn test to identify the significant pairs
signif_cancer_irt_prm <- krus_cancer_irt$parameter[krus_cancer_irt$p_value < 0.05]

for (var in signif_cancer_irt_prm) {
  print(var)
  dunn.test(train[[var]], train$CANCERTYPE, method = "holm")
}
```

### 3.2.1 BASEGI (p<0.05)
```{r}
# description value of each group
train %>%
  group_by(CANCERTYPE) %>%
  summarise(median = median(BASEGI))

# ggplot
p1 <- ggboxplot(train, x = "CANCERTYPE", y = "BASEGI",
          fill = "CANCERTYPE", palette = "nejm",
          add = "jitter", add.params = list(alpha = 0.5)) +
  geom_signif(
    comparisons = list(c("Genitourinary", "Breast"), c("Genitourinary", "Gastrointestinal")),  
    annotations = c("p = 0.013", "p = 0.002"),       
    y_position = c(5.6, 6),        
    tip_length = 0.02, textsize = 6.2
  )+
  stat_compare_means(method = "kruskal.test",
                     label.x = 0.8,
                     label.y = max(train$BASEGI)*1.1,
                     size = 6.2) +
  labs(x = "Cancer Type",y = "Baseline PRO") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )
print(p1)

ggsave("plot/figure6b.tiff",
       plot = p1,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)

```

### 3.2.1 BASEPD (p<0.05)
```{r}
# description value of each group
train %>%
  group_by(CANCERTYPE) %>%
  summarise(median = median(BASEPD))

# ggplot
p2 <- ggboxplot(train, x = "CANCERTYPE", y = "BASEPD",
          fill = "CANCERTYPE", palette = "nejm",
          add = "jitter", add.params = list(alpha = 0.5)) +
  geom_signif(
    comparisons = list(c("Thoracic", "Gastrointestinal")),  
    annotations = c("p = 0.017"),       
    y_position = 5.9,        
    tip_length = 0.02,textsize = 6.2
  )+
  stat_compare_means(method = "kruskal.test",
                     label.x = 0.8,
                     label.y = 7, size = 6.2) +
  labs(x = "Cancer Type", y = "Baseline PRO") +
  scale_y_continuous(limits = c(-1, 7), breaks = seq(-1, 7, 2)) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )

print(p2)

ggsave("plot/figure6c.tiff",
       plot = p2,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)

```



## 3.2.2 Previous treatment

```{r}
krus_iv_irt <- data.frame()
for (var in vars) {
  kw <- kruskal.test(as.formula(paste(var, "~ IV")), data = train)
  krus_iv_irt <- rbind(krus_iv_irt, data.frame(
    parameter = var,
    H = round(kw$statistic, 3),
    p_value = round(kw$p.value, 4)
  ))
}

print(krus_iv_irt)
#parameter      H p_value
#Kruskal-Wallis chi-squared	BASEGI	18.584	0.0009	
#Kruskal-Wallis chi-squared7	SLPPS	18.100	0.0012	
#Kruskal-Wallis chi-squared11	SLPPD	22.441	0.0002	

# Dunn test to identify the significant pairs
signif_iv_irt_prm <- krus_iv_irt$parameter[krus_iv_irt$p_value < 0.05]

for (var in signif_iv_irt_prm) {
  print(var)
  dunn.test(train[[var]], train$IV, method = "holm")
}

table(train$IV)
```
### 3.2.2 SLPPS (p<0.05)
```{r}
train %>%
  group_by(IV) %>%
  summarise(median = median(SLPPS))

# ggplot
p3 <- ggboxplot(train, x = "IV", y = "SLPPS",
          fill = "IV", palette = "nejm",
          add = "jitter", add.params = list(alpha = 0.5)) +
  geom_signif(comparisons = list(c("Target", "Chemo")),  
              annotations = c("p = 0.009"), 
              y_position = 6, tip_length = 0.004,textsize = 6.2) +
  stat_compare_means(method = "kruskal.test",
                     label.x = 2,label.y = 9.7, size = 6.2) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0, 10, 2.5)) +
  labs(x = "Treatment", y = "SLP") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )
print(p3)

ggsave("plot/figure6e.tiff",
       plot = p3,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)
```

### 3.2.3 SLPPD (p<0.05)

```{r}
train %>%
  group_by(IV) %>%
  summarise(median = median(SLPPD))

# ggplot
p4 <- ggboxplot(train, x = "IV", y = "SLPPD",
          fill = "IV", palette = "nejm",
          add = "jitter", add.params = list(alpha = 0.5)) +
  geom_signif(comparisons = list(c("Target", "Chemo"),c("None or Unclassified", "Chemo")),  
              annotations = c("p = 0.001", "p = 0.002"),    
              y_position = c(10,13), tip_length = 0.004,textsize = 6.2) +
  stat_compare_means(method = "kruskal.test",
                     label.x = 0.8,
                     label.y = 19.8, size = 6.2) +
  scale_y_continuous(limits = c(0,20)) +
  labs(x = "Treatment",y = "SLP") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )

print(p4)

ggsave("plot/figure6f.tiff",
       plot = p4,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)

```

# 3.3 PRO total model comparsion
## 3.3.1 Cancer type
```{r}
train_pro <- train_demo %>%
        left_join(train_pro_prm,by = "PTID")

vars_pro <- c("BASEP","Emax","Kd","SLP")

# Kruskal-Wallis results
krus_cancer_pro <- data.frame()
for (var in vars_pro) {
  kw <- kruskal.test(as.formula(paste(var, "~ CANCERTYPE")), data = train_pro)
  krus_cancer_pro <- rbind(krus_cancer_pro, data.frame(
    parameter = var,
    H = round(kw$statistic, 3),
    p_value = round(kw$p.value, 4)
  ))
}

print(krus_cancer_pro)

#parameter      H p_value (no significant difference)
#Kruskal-Wallis chi-squared	BASEP	8.107	0.1504	
#Kruskal-Wallis chi-squared1	Emax	10.934	0.0527	
#Kruskal-Wallis chi-squared2	Kd	10.622	0.0594	
#Kruskal-Wallis chi-squared3	SLP	3.806	0.5777

```

Generate the plot for BASEP (total baseline) and compare with IRT individual baseline
```{r}
# --- paramter BASEP 
# description value of each group
train_pro %>%
  group_by(CANCERTYPE) %>%
  summarise(median = median(BASEP))

# ggplot
pro_p1 <- ggboxplot(train_pro, x = "CANCERTYPE", y = "BASEP",
          fill = "CANCERTYPE", palette = "nejm",
          add = "jitter", add.params = list(alpha = 0.5)) +
  stat_compare_means(method = "kruskal.test",
                     label.x = 0.8,
                     label.y = 0.9, size = 6.2) +
  labs(x = "Cancer Type", y = "Baseline PRO") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )

print(pro_p1)

ggsave("plot/figure6a.tiff",
       plot = pro_p1,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)
```

## 3.3.2 Previou Treatment
```{r}
krus_iv_pro <- data.frame()
for (var in vars_pro) {
  kw <- kruskal.test(as.formula(paste(var, "~ IV")), data = train_pro)
  krus_iv_pro <- rbind(krus_iv_pro, data.frame(
    parameter = var,
    H = round(kw$statistic, 3),
    p_value = round(kw$p.value, 4)
  ))
}

print(krus_iv_pro)

#parameter      H p_value (no significant difference)
#Kruskal-Wallis chi-squared	BASEP	6.686	0.1534	
#Kruskal-Wallis chi-squared1	Emax	5.043	0.2830	
#Kruskal-Wallis chi-squared2	Kd	1.418	0.8411	
#Kruskal-Wallis chi-squared3	SLP	6.509	0.1642	

```

Generate the plot for SLP (total deterioration rate) and compare with IRT individual deterioration rate

```{r}
# --- parameter SLP 
# description value of each group
train_pro %>%
  group_by(CANCERTYPE) %>%
  summarise(median = median(SLP))

# ggplot
pro_p2 <- ggboxplot(train_pro, x = "IV", y = "SLP",
                    fill = "IV", palette = "nejm",
                    add = "jitter", add.params = list(alpha = 0.5)) +
  stat_compare_means(method = "kruskal.test",
                     label.x = 0.8,
                     label.y = 12.3, size = 6.2) +
  labs(x = "Treatment", y = "SLP") +
  scale_y_continuous(limits = c(-7.5,12.5), breaks = seq(-7.5,12.5,5)) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 18.5, color = "black", hjust =1, angle = 30),
        axis.text.y = element_text(size = 18, color = "black"),
        axis.title = element_text(size = 19, face = "bold")
  )

print(pro_p2)

ggsave("plot/figure6d.tiff",
       plot = pro_p2,
       width = 10.5,
       height = 6,
       units = "in",
       dpi = 300)
```


















