---
title: "Data exploratory - Individual Score"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggsci)
library(survival)
library(survminer)
library(gt)
library(gtsummary)
library(forestplot)
library(broom)
library(car)

```

# 2. Import the raw data 
After exploratory analysis, level 3 and level 4 were recorded and combined together for further IRT model (limited level of level 4 data for all symptoms)

```{r, echo=FALSE}
# full PRO data
pro <- read.csv(here("data","fulldata_IRT.csv")) %>%
        mutate(DV = if_else(DV == 4, 3, DV),DV2 = 4 - DV) %>%
        # re-code the binary outcome variable
        mutate(DISCONTINUE = case_when(is.na(DISCONTINUE) ~ "0",TRUE ~ "1")) %>%
        # rearrange the columns
        select(PTID,CONSENT_DT,PROCTCAE,VAR,TIME,TAFR,DV,DV2,everything())

# Time to Event (no time for discontinuation data)
endpoint <- read.csv(here("data","endpoint.csv")) %>% 
        mutate(DIS = case_when(is.na(C12_discontinuing_overall) ~ "0",TRUE ~ "1")) %>%
        # missing means survival, recoded to 0 - survival; 1 - death
        mutate(C9_death = case_when(is.na(C9_death) ~ "0",TRUE ~ "1"))
```

# 3.1. Data exploratory

## 3.1.1 PRO level distribution of full data and baseline

```{r, echo=FALSE}
#--------------------------------------
# Full data set distribution
#--------------------------------------
pro.count <- pro %>%
  mutate(DV = if_else(is.na(DV), "Missing", as.character(DV)),
         DV = factor(DV, levels = c("Missing","0","1","2","3"))
        ) %>%
  count(PROCTCAE, DV, name = "n") 

PRO_map <- c(PROCTCAE_001 = "Activity",PROCTCAE_002 = "Depression",PROCTCAE_003 = "Appetite",
             PROCTCAE_004 = "Pain_frequency",PROCTCAE_005 = "Pain_severity",PROCTCAE_006 = "Pain_interference",
             PROCTCAE_007 = "Nausea_frequency",PROCTCAE_008 = "Nausea_severity",PROCTCAE_009 = "Vomit_frequency",
             PROCTCAE_010 = "Diarrhea_frequency",PROCTCAE_011 = "Constipation_severity",PROCTCAE_012 = "Dyspnea_severity",
             PROCTCAE_013 = "Dyspnea_interference", PROCTCAE_014 = "Insomnia_severity")

pro.color <- c("Missing" = "grey50","0" = "darkgreen",  "1" = "green3", 
               "2" = "orange", "3" = "#DB5829")

pro.count.plot <- pro.count %>%
    mutate(PROCTCAE = case_when(
    grepl(names(PRO_map)[1], PROCTCAE, ignore.case = TRUE) ~ PRO_map[1],
    grepl(names(PRO_map)[2], PROCTCAE, ignore.case = TRUE) ~ PRO_map[2],
    grepl(names(PRO_map)[3], PROCTCAE, ignore.case = TRUE) ~ PRO_map[3],
    grepl(names(PRO_map)[4], PROCTCAE, ignore.case = TRUE) ~ PRO_map[4],
    grepl(names(PRO_map)[5], PROCTCAE, ignore.case = TRUE) ~ PRO_map[5],
    grepl(names(PRO_map)[6], PROCTCAE, ignore.case = TRUE) ~ PRO_map[6],
    grepl(names(PRO_map)[7], PROCTCAE, ignore.case = TRUE) ~ PRO_map[7],
    grepl(names(PRO_map)[8], PROCTCAE, ignore.case = TRUE) ~ PRO_map[8],
    grepl(names(PRO_map)[9], PROCTCAE, ignore.case = TRUE) ~ PRO_map[9],
    grepl(names(PRO_map)[10], PROCTCAE, ignore.case = TRUE) ~ PRO_map[10],
    grepl(names(PRO_map)[11], PROCTCAE, ignore.case = TRUE) ~ PRO_map[11],
    grepl(names(PRO_map)[12], PROCTCAE, ignore.case = TRUE) ~ PRO_map[12],
    grepl(names(PRO_map)[13], PROCTCAE, ignore.case = TRUE) ~ PRO_map[13],
    grepl(names(PRO_map)[14], PROCTCAE, ignore.case = TRUE) ~ PRO_map[14]
  ))


p1 <- ggplot(pro.count.plot, aes(x = PROCTCAE, y = n, fill = DV)) +
  geom_col(width = 0.8, color = "white", linewidth = 0.2) +
  scale_fill_manual(values = pro.color, name = "Score",
                labels = c("3" = "3/4", "2" = "2", "1" = "1", "0" = "0")) +
  labs(x = NULL, y = "Count") +
  theme_bw() +
  theme(
    legend.position = "top",
    legend.text  = element_text(size = 15),
    legend.title  = element_text(size = 15),
    axis.text.x = element_text(angle = 45, hjust = 1, color = "black",size = 16),
    axis.text.y = element_text(color = "black",size = 16),
    axis.title = element_text(size= 15, face = "bold"),
  )

print(p1)

ggsave("plot/Supplement1a.tiff",
       plot = p1,
       width = 14,
       height = 6,
       units = "in",
       dpi = 300)

#--------------------------------------
# Baseline Distribution
#--------------------------------------
pro.count2 <- pro %>% 
  filter(TAFR == "0") %>%
  mutate(
    DV = if_else(is.na(DV), "Missing", as.character(DV)),
    DV = factor(DV, levels = c("Missing","0","1","2","3"))
  ) %>%
  count(PROCTCAE, DV, name = "n") 

ggplot(pro.count2, aes(x = PROCTCAE, y = n, fill = DV)) +
  geom_col(width = 0.8, color = "white", linewidth = 0.2) +
  scale_fill_manual(values = pro.color, name = "Score") +
  labs(x = NULL, y = "Count") +
  theme_bw() +
  theme(
    legend.position = "top",
    legend.text  = element_text(size = 15),
    legend.title  = element_text(size = 15),
    axis.text.x = element_text(angle = 45, hjust = 1, color = "black",size = 16),
    axis.text.y = element_text(color = "black",size = 16),
    axis.title = element_text(size= 15, face = "bold"),
  )

```

## 3.1.2 Time to Event data analysis

```{r, include=FALSE, echo=FALSE}

# exclude patients without PRO data (these data didnt include in the model)
PTID_no_missing <- pro %>%
        filter(is.na(DV) == FALSE) %>%
        pull(PTID) %>%
        unique()

endpoint <- endpoint %>%
        filter(PTID %in% PTID_no_missing)

# Occurrence of each event
table(endpoint$C9_death)
table(endpoint$C12_dose_reduction)
table(endpoint$C12_er)
table(endpoint$C12_hospital)
table(endpoint$DIS)

#--------------------------------------
# Time to event distribution 
#--------------------------------------
#-- Dose reduction
red_time <- endpoint %>%  
  pivot_longer(starts_with("C12_DOSE_REDUCTION_DAY"),names_to = "reduction",values_to = "time") %>%
  filter(!is.na(time),time >= 0) %>%
  group_by(PTID) %>%
  summarise(time = min(time), .groups = "drop") %>%
  select(PTID, time)
summary(red_time$time)

#-- er visit
er_time <- endpoint %>%  
  pivot_longer(starts_with("C12_er_DAY"),names_to = "er",values_to = "time") %>%
  filter(!is.na(time),time >= 0) %>%
  group_by(PTID) %>%
  summarise(time = min(time), .groups = "drop") %>%
  select(PTID, time)
summary(er_time$time)

#-- Hospitalization
hospital_time <- endpoint %>%  
  pivot_longer(starts_with("C12_hospital_DAY"),names_to = "hospital",values_to = "time") %>%
  filter(!is.na(time),time >= 0) %>%
  group_by(PTID) %>%
  summarise(time = min(time), .groups = "drop") %>%
  select(PTID, time)
summary(hospital_time$time)

#-- Death
survival_time <- endpoint %>%
  rename(time = C9_death_date) %>%
  filter(!is.na(time)) %>%
  group_by(PTID) %>%
  summarise(time = min(time), .groups = "drop") %>%
  select(PTID, time)
summary(survival_time$time)  

#--------------------------------------
# K-M Curve
#--------------------------------------
km_plot <- pro %>%
  filter(PTID %in% PTID_no_missing) %>%  #filter out patients with no PRO data
  rename(DTH = DEATH,DTHDY = DEATH.DATE) %>%
  mutate(DTH = as.numeric(case_when(is.na(DTH) ~ "0",TRUE ~ "1"))) %>%
  select(PTID,TIME,DV,DTH,DTHDY)

# generate OS data frame
os_row_list <- list()

for (id in unique(km_plot$PTID)) {
  patient_data <- km_plot[km_plot$PTID == id, ]
  template <- km_plot[km_plot$PTID == id, ][1, ]
  
  # LAST observation (TIME=DTHDY, QSFLAG="OS")
  if (!is.na(template$DTHDY)) {
    # If DTH happened: use DTHDY
    endpoint <- template
    endpoint$TIME <- template$DTHDY
    endpoint$DV <- template$DTH

    os_row_list[[length(os_row_list) + 1]] <- endpoint
  } else {
    # If DTH not happen but censor: Use last observation TIME
    last_obs_time <- max(patient_data$TIME, na.rm = TRUE)
    
    endpoint <- template
    endpoint$TIME <- last_obs_time
    endpoint$DV <- 0

    os_row_list[[length(os_row_list) + 1]] <- endpoint
  }
}

km_final <- bind_rows(os_row_list) %>%
          select(-DTH,-DTHDY) %>%
          rename(DTH = DV, DTHDY = TIME) %>%
          mutate(DTHDY = DTHDY)

# KM fit and curve
km_fit <- survfit(Surv(DTHDY, DTH) ~ 1, data = km_final)

km <- ggsurvplot(
  km_fit,
  data = km_final,
  legend = "none", 
  conf.int = FALSE,
  palette = "jco",
  
  xlab = "Time (days)",
  xlim = c(0, 720),
  break.time.by = 120,

  ylab = "Overall Survival (%)",
  ylim = c(0, 1),              
  yticklabs = c("0%", "25%","50%", "75%", "100%"),
  surv.scale = "percent",
  
  size = 1.2,
  censor = TRUE,
  censor.shape = "+",
  censor.size = 6,
  
  # font size
  font.x = c(15, "bold", "black"),      
  font.y = c(15, "bold", "black"),      
  font.tickslab = c(15, "plain","black"),        
  
  risk.table = TRUE,
  risk.table.height = 0.22,
  risk.table.fontsize = 5,     
  tables.theme = theme_cleantable() +
    theme(axis.text.x = element_text(size = 15, color = "black"),  
          axis.text.y = element_text(size = 15, color = "black"),
          axis.title.y = element_text(size = 15, color = "black", face = "bold"), 
          axis.title.x = element_text(size = 15, color = "black", face = "bold")),
  
  ggtheme = theme_bw()
)

print(km)

tiff("plot/figure1e_KM.tiff",width = 7, height = 6, units = "in", res = 300)
print(km)
dev.off()

```

## 3.1.3 Missingness check
```{r, include=FALSE, echo=FALSE}
# filter the first 360 days data
pro_360 <- pro %>%
        filter(TIME <= 360)

sum(is.na(pro_360$DV))/length(pro_360$DV) #23.8%

pro_item <- pro_360 %>%
      group_by(PROCTCAE) %>%
      summarise(
        Total = n(),
        Missing = sum(is.na(DV)),
        NA_percent = Missing/Total,
      ) %>%
      ungroup() %>%
      mutate(pct = NA_percent * 100)

# cut time into several windows (60 days as a window)
breaks <- c(seq(0, 360, by = 60))
win_labs <- as.character(seq(60, by = 60, length.out = length(breaks) - 1))

pro.na <- pro_360 %>%
      mutate(window = cut(TIME, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
      group_by(PROCTCAE,window) %>%
      summarise(
        Total = n(),
        Missing = sum(is.na(DV)),
        NA_percent = Missing/Total,
      ) %>%
      ungroup() %>%
      mutate(pct = NA_percent * 100) %>%
      group_by(PROCTCAE) %>%
      # for later plot setting
      mutate(vjust_val = ifelse(row_number() %% 2 == 0, 1.2, -0.5)) %>%
      ungroup()

# rename the PROCTCAE to original names
pro.na.plot <- pro.na %>%
          mutate(PROCTCAE = case_when(
          grepl(names(PRO_map)[1], PROCTCAE, ignore.case = TRUE) ~ PRO_map[1],
          grepl(names(PRO_map)[2], PROCTCAE, ignore.case = TRUE) ~ PRO_map[2],
          grepl(names(PRO_map)[3], PROCTCAE, ignore.case = TRUE) ~ PRO_map[3],
          grepl(names(PRO_map)[4], PROCTCAE, ignore.case = TRUE) ~ PRO_map[4],
          grepl(names(PRO_map)[5], PROCTCAE, ignore.case = TRUE) ~ PRO_map[5],
          grepl(names(PRO_map)[6], PROCTCAE, ignore.case = TRUE) ~ PRO_map[6],
          grepl(names(PRO_map)[7], PROCTCAE, ignore.case = TRUE) ~ PRO_map[7],
          grepl(names(PRO_map)[8], PROCTCAE, ignore.case = TRUE) ~ PRO_map[8],
          grepl(names(PRO_map)[9], PROCTCAE, ignore.case = TRUE) ~ PRO_map[9],
          grepl(names(PRO_map)[10], PROCTCAE, ignore.case = TRUE) ~ PRO_map[10],
          grepl(names(PRO_map)[11], PROCTCAE, ignore.case = TRUE) ~ PRO_map[11],
          grepl(names(PRO_map)[12], PROCTCAE, ignore.case = TRUE) ~ PRO_map[12],
          grepl(names(PRO_map)[13], PROCTCAE, ignore.case = TRUE) ~ PRO_map[13],
          grepl(names(PRO_map)[14], PROCTCAE, ignore.case = TRUE) ~ PRO_map[14]
          ))

pro.na.plot <- pro.na.plot %>%
  mutate(PROCTCAE = factor(PROCTCAE, levels = c("Activity", "Depression", "Appetite",
                                              "Pain_frequency","Pain_severity","Pain_interference","Nausea_frequency",
                                              "Nausea_severity","Vomit_frequency","Diarrhea_frequency","Constipation_severity",
                                              "Dyspnea_severity","Dyspnea_interference","Insomnia_severity")))

p2 <- ggplot(pro.na.plot, aes(x = window, y = pct, group = PROCTCAE)) +
  geom_line(size = 1, color = "#1964B0", alpha = 1) +  
  geom_point(size = 1.5, color = "#1964B0", alpha = 1) +  
  geom_text(aes(label = sprintf("%.1f%%", NA_percent * 100), 
                vjust = vjust_val), size = 4) +
  facet_wrap(~PROCTCAE, ncol = 5) +  
  scale_y_continuous(labels = function(x) paste0(x, "%"), 
                     expand = expansion(mult = c(0, 0.15)),
                     breaks = seq(0,100,20), limits = c(0,100)) +  
  labs(
    x = "Time (days)",
    y = "Missing Percentage"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 15),
    strip.background = element_rect(fill = "lightgray"),
    axis.text.x = element_text(color = "black",size = 15),
    axis.text.y = element_text(color = "black",size = 15),
    axis.title = element_text(size = 15, face = "bold")
  )
print(p2)

ggsave("plot/supplement1b.tiff",
       plot = p2,
       width = 14,
       height = 6,
       units = "in",
       dpi = 300)
```

## 3.1.4 PRO Score change over time
```{r, include=FALSE, echo=FALSE}
#--------------------------------------
# Individual Symptom Score
#--------------------------------------
# remove missing rows and recalculate the first time (since the original first row may be removed) and TAFR
no.missing <- pro %>%
  filter(is.na(DV) == FALSE) %>%
  group_by(PTID) %>%
  mutate(FIRSTTIME = first(TIME)) %>%
  mutate(TAFR = TIME - FIRSTTIME) %>%
  ungroup() 

# calculate the pro mean and sd for each window and generate the plots
breaks2 <- c(seq(0, 360, by = 60), 400)     # 30 days as a window, last 40 days as a window
win_labs2 <- c(seq(60, 360, by = 60), 400) # W01..W13

pro.mean <- no.missing %>%
  mutate(window = cut(TAFR, breaks = breaks2, include.lowest = TRUE, labels = win_labs2)) %>%
  group_by(PROCTCAE, window) %>%
  summarise(
    count = n(),
    mean = mean(DV, na.rm = TRUE),
    sd = sd(DV, na.rm = TRUE),
    se = sd/sqrt(count),
    .groups = "drop"
  )

pro.mean <- pro.mean %>%
  mutate(PROCTCAE = case_when(
    grepl(names(PRO_map)[1], PROCTCAE, ignore.case = TRUE) ~ PRO_map[1],
    grepl(names(PRO_map)[2], PROCTCAE, ignore.case = TRUE) ~ PRO_map[2],
    grepl(names(PRO_map)[3], PROCTCAE, ignore.case = TRUE) ~ PRO_map[3],
    grepl(names(PRO_map)[4], PROCTCAE, ignore.case = TRUE) ~ PRO_map[4],
    grepl(names(PRO_map)[5], PROCTCAE, ignore.case = TRUE) ~ PRO_map[5],
    grepl(names(PRO_map)[6], PROCTCAE, ignore.case = TRUE) ~ PRO_map[6],
    grepl(names(PRO_map)[7], PROCTCAE, ignore.case = TRUE) ~ PRO_map[7],
    grepl(names(PRO_map)[8], PROCTCAE, ignore.case = TRUE) ~ PRO_map[8],
    grepl(names(PRO_map)[9], PROCTCAE, ignore.case = TRUE) ~ PRO_map[9],
    grepl(names(PRO_map)[10], PROCTCAE, ignore.case = TRUE) ~ PRO_map[10],
    grepl(names(PRO_map)[11], PROCTCAE, ignore.case = TRUE) ~ PRO_map[11],
    grepl(names(PRO_map)[12], PROCTCAE, ignore.case = TRUE) ~ PRO_map[12],
    grepl(names(PRO_map)[13], PROCTCAE, ignore.case = TRUE) ~ PRO_map[13],
    grepl(names(PRO_map)[14], PROCTCAE, ignore.case = TRUE) ~ PRO_map[14]
  ))

pro.mean <- pro.mean %>%
            mutate(PROCTCAE = factor(PROCTCAE, levels = c("Activity", "Depression", "Appetite",
                              "Pain_frequency","Pain_severity","Pain_interference","Nausea_frequency",
                              "Nausea_severity","Vomit_frequency","Diarrhea_frequency","Constipation_severity",
                              "Dyspnea_severity","Dyspnea_interference","Insomnia_severity")))


p3 <- ggplot(pro.mean, aes(x = window, y = mean)) +
  geom_line(aes(group = 1), size = 1,color = "#1964B0") +
  geom_point(size = 1,color = "#1964B0") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),width = 0.3,linewidth = 1,color = "#1964B0") +
  #scale_fill_jco() +
  facet_wrap(~PROCTCAE, ncol = 5) +
  scale_y_continuous(breaks = seq(-1,3,1), limits = c(-1,3)) +
  labs(
    x = "Time (days)",
    y = "Score (mean ± SD)"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 15),
    axis.text.x = element_text(color = "black",size = 15, hjust = 0.5),
    axis.text.y = element_text(color = "black",size = 15),
    strip.background = element_rect(fill = "lightgray"),
    axis.title = element_text(size = 15, face = "bold")
  )

print(p3)

ggsave("plot/figure1b.tiff",
       plot = p3,
       width = 16,
       height = 6,
       units = "in",
       dpi = 300)

```

## 3.1.5 demographic table
```{r, include=FALSE, echo=FALSE}

#--------------------------------------
# Demographic Table
#--------------------------------------
# convert some variables and regroup the IV and PO treatment
no.missing <- no.missing %>%
        mutate(DEATH = case_when(is.na(DEATH) ~ "0",TRUE ~ "1")) %>%
        # regroup the IV and PO treatment groups 
        mutate(REGIMEN_IV = case_when(    
         grepl("Chemotherapy only",REGIMEN_IV,ignore.case = TRUE) ~ "1",
         grepl("Targeted only",REGIMEN_IV,ignore.case = TRUE) ~ "2",
         grepl("Immunotherapy only",REGIMEN_IV,ignore.case = TRUE) ~ "3",
         REGIMEN_IV %in% c("Chemo_Targeted","three combine","Chemo_Immunotherapy",
                      "Immuno_Targeted") ~ "4",
         TRUE ~"5")) %>%
  
        mutate(REGIMEN_PO = case_when(
         grepl("TKI",REGIMEN_PO,ignore.case = TRUE) ~ "1",
         TRUE ~"2")) %>%
  
        mutate(across(c(GENDER, ETHNICITY, RACE, CANCERTYPE, ECOG,REGIMEN_IV, REGIMEN_PO, C12_DOSE_REDUCTION,
                DISCONTINUE,PROCTCAE, VAR, ER, HOSPITAL,REGIMEN_IV,REGIMEN_PO),as.factor)) %>%
        select(-CONSENT_DT) %>%

        rename(SEX = GENDER, IV = REGIMEN_IV, PO = REGIMEN_PO, SURVIVAL_TIME = DEATH.DATE) %>%
        select(PTID,PROCTCAE,VAR,TIME,TAFR,DV,DV2,MDV,ER,HOSPITAL,C12_DOSE_REDUCTION,DISCONTINUE,DEATH,SURVIVAL_TIME,
               everything())

demo.table <- no.missing %>%
  group_by(PTID) %>%
  slice(1) %>%
  ungroup() %>%
  select(AGE,SEX,ETHNICITY,RACE,HEIGHT,CANCERTYPE,ECOG,IV,PO) %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{median} ({min}, {max})",  
      all_categorical() ~ "{n} ({p}%)"              
    ),
    digits = list(
      all_continuous() ~ 1,
      all_categorical() ~ 1
    )
  ) %>%
  modify_header(label ~ "**Variable**", stat_0 ~ "**Summary (N = {N})**")

print(demo.table)

```

## 3.1.6 multivariable logistic regression
### 3.1.6.1 Regroup the variables
```{r, include=FALSE, echo=FALSE}
# Selected covariate: sex,race,cancer type,ecog,iv and po treatment
demo.clean <- no.missing %>%
          group_by(PTID) %>%
          # one row for each patient
          slice(1) %>%
          ungroup() %>%
  # combine Asian,native Hawaiian,Multiple Races Selected and NA to others.
        mutate(RACE = factor(case_when(
          RACE == 1|RACE == 2 | RACE == 4 | RACE == 99 | is.na(RACE) ~ "Other",
          RACE == 3 ~ "Black",
          RACE == 5 ~ "White"),levels = c("White", "Black", "Other"))) %>%
  # rename sex to male and female
        mutate(SEX = factor(if_else(SEX == 1, "Male", "Female"))) %>%
  # combine ECOG 2 and 3 together
        mutate(ECOG = factor(if_else(ECOG == "3", "2", as.character(ECOG)))) %>%
  # rename cancer type
        mutate(CANCERTYPE = factor(case_when(
                CANCERTYPE == 1 ~ "Thoracic",
                CANCERTYPE == 2 ~ "Breast",
        #(Colorectal/Anal,Gastro-esophageal,Small bowel and 'Pancreas, Hepatobiliary)
                CANCERTYPE %in% c(3,9,10) ~ "Gastrointestinal", 
                CANCERTYPE %in% c(4,8) ~ "Genitourinary",
                CANCERTYPE == 5 ~ "Gynecologic",
                CANCERTYPE %in% c(6,7,99) ~ "Other tumors")),
                CANCERTYPE = relevel(CANCERTYPE, ref = "Thoracic")) %>%
  # rename IV treatment
        mutate(IV = factor(case_when(
          IV == 1 ~ "Chemo",
          IV == 2 ~ "Target",
          IV == 3 ~ "Immuno",
          IV == 4 ~ "Combination",
          IV == 5 ~ "None or Unclassified")),
          IV = relevel(IV, ref = "Chemo")) %>%
  # rename PO treatment
        mutate(PO = factor(case_when(
          PO == 1 ~ "TKI",
          PO == 2 ~ "None or Unclassified")),
          PO = relevel(PO, ref = "None or Unclassified"))

```

### 3.1.6.2 Dose reduction forest plot
```{r, include=FALSE, echo=FALSE}
logis.red <- glm(C12_DOSE_REDUCTION ~ SEX + RACE + CANCERTYPE + ECOG + IV + PO,
             data = demo.clean,
             family = binomial)

summary(logis.red)

result.red <- tidy(logis.red, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR = estimate,
    CI_lower = conf.low,
    CI_upper = conf.high,
    OR_CI = sprintf("%.2f (%.2f-%.2f)", OR, CI_lower, CI_upper),
    p_formatted = case_when(
      p.value < 0.001 ~ "<0.001",
      p.value < 0.01 ~ sprintf("%.3f", p.value),
      TRUE ~ sprintf("%.2f", p.value)
    ),
    sig_mark = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    Variable = case_when(
      grepl("SEX", term) ~ paste0("Sex: ", gsub("SEX", "", term), " vs. Female"),
      grepl("RACE", term) ~ paste0("Race: ", gsub("RACE", "", term), " vs. White"),
      grepl("CANCERTYPE", term) ~ paste0("Cancer: ", gsub("CANCERTYPE", "", term), " vs. Thoracic"),
      grepl("ECOG", term) ~ paste0("ECOG: ", ifelse(gsub("ECOG", "", term) == "2", "2/3", gsub("ECOG", "", term)), " vs. ECOG 0"),
      grepl("IV", term) ~ paste0("IV: ", gsub("IV", "", term), " vs. Chemo"),
      grepl("PO", term) ~ paste0("PO: ", gsub("PO", "", term), " vs. None or Unclassified"),
      TRUE ~ term
    )
  )

# Make Forest plot
tabletext <- cbind(
  c("Variable", result.red$Variable),
  c("OR (95% CI)", result.red$OR_CI),
  c("P Value", paste0(result.red$p_formatted, result.red$sig_mark))
)

tiff("plot/supplement2a.tiff", 
     width = 6, height = 4.3, units = "in", res = 300)

forestplot(
  labeltext = tabletext,
  mean = c(NA, result.red$OR),
  lower = c(NA, result.red$CI_lower),
  upper = c(NA, result.red$CI_upper),
  is.summary = c(TRUE, rep(FALSE, nrow(result.red))),
  align = c("l", "l", "c"),

  # X-axis format
  xlog = TRUE,
  xticks = c(0.1,1,10),
  xticks.digits = 1,
  xlab = "Odds Ratio (OR)",
  
  # Color
  col = fpColors(
    box = "black",
    line = "black" ,
    zero = "grey50"),
  
  # Line and Point format
  lwd.xaxis = 1.5,
  lwd.ci = 1.5,
  lwd.zero = 1, #OR = 1 line
  boxsize = 0.25,
  vertices = TRUE,
  
  # Font Format
  txt_gp = fpTxtGp(
    label = gpar(fontsize = 10),
    ticks = gpar(fontsize = 20),
    xlab = gpar(fontsize = 16,fontface = "bold"),
    title = gpar(
      fontsize = 13,           
      fontface = "bold",   
      lineheight = 10)),
  
  # Layout setting
  graphwidth = unit(1.5, "inches"),
  colgap = unit(2, "mm"),
  lineheight = unit(6, "mm")
)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.08, 
  gp = gpar(cex = 0.7, fontface = "italic")
)

dev.off()


```
### 3.1.6.3 ER Visit forest plot
```{r, include=FALSE, echo=FALSE}
logis.er <- glm(ER ~ SEX + RACE + CANCERTYPE + ECOG + IV + PO,
                 data = demo.clean,
                 family = binomial)

summary(logis.er)

result.er <- tidy(logis.er, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR = estimate,
    CI_lower = conf.low,
    CI_upper = conf.high,
    OR_CI = sprintf("%.2f (%.2f-%.2f)", OR, CI_lower, CI_upper),
    p_formatted = case_when(
      p.value < 0.001 ~ "<0.001",
      p.value < 0.01 ~ sprintf("%.3f", p.value),
      TRUE ~ sprintf("%.2f", p.value)
    ),
    sig_mark = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    Variable = case_when(
      grepl("SEX", term) ~ paste0("Sex: ", gsub("SEX", "", term), " vs. Female"),
      grepl("RACE", term) ~ paste0("Race: ", gsub("RACE", "", term), " vs. White"),
      grepl("CANCERTYPE", term) ~ paste0("Cancer: ", gsub("CANCERTYPE", "", term), " vs. Thoracic"),
      grepl("ECOG", term) ~ paste0("ECOG: ", ifelse(gsub("ECOG", "", term) == "2", "2/3", gsub("ECOG", "", term)), " vs. ECOG 0"),
      grepl("IV", term) ~ paste0("IV: ", gsub("IV", "", term), " vs. Chemo"),
      grepl("PO", term) ~ paste0("PO: ", gsub("PO", "", term), " vs. None or Unclassified"),
      TRUE ~ term
    )
  )

# Make Forest plot
tabletext <- cbind(
  c("Variable", result.er$Variable),
  c("OR (95% CI)", result.er$OR_CI),
  c("P Value", paste0(result.er$p_formatted, result.er$sig_mark))
)

tiff("plot/supplement2b.tiff", 
     width = 6, height = 4.3, units = "in", res = 300)

forestplot(
  labeltext = tabletext,
  mean = c(NA, result.er$OR),
  lower = c(NA, result.er$CI_lower),
  upper = c(NA, result.er$CI_upper),
  is.summary = c(TRUE, rep(FALSE, nrow(result.er))),
  align = c("l","l","c"),

  # X-axis format
  xlog = TRUE,
  xticks = c(0.1,1,10),
  xticks.digits = 1,
  xlab = "Odds Ratio (OR)",
  
  # Color
  col = fpColors(
    box = "black",
    line = "black" ,
    zero = "grey50"),
  
  # Line and Point format
  lwd.xaxis = 1.5,
  lwd.ci = 1.5,
  lwd.zero = 1, #OR = 1 line
  boxsize = 0.25,
  vertices = TRUE,
  
  # Font Format
  txt_gp = fpTxtGp(
    label = gpar(fontsize = 10),
    ticks = gpar(fontsize = 20),
    xlab = gpar(fontsize = 16,fontface = "bold"),
    title = gpar(
      fontsize = 13,           
      fontface = "bold",   
      lineheight = 10)),
  
  # Layout setting
  graphwidth = unit(1.5, "inches"),
  colgap = unit(2, "mm"),
  lineheight = unit(6, "mm")
)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.08, 
  gp = gpar(cex = 0.7, fontface = "italic")
)

dev.off()

```
### 3.1.6.4 Dose discontinuation forest plot
```{r, include=FALSE, echo=FALSE}
logis.dis <- glm(DISCONTINUE ~ SEX + RACE + CANCERTYPE + ECOG + IV + PO,
                data = demo.clean,
                family = binomial)

summary(logis.dis)

result.dis <- tidy(logis.dis, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR = estimate,
    CI_lower = conf.low,
    CI_upper = conf.high,
    OR_CI = sprintf("%.2f (%.2f-%.2f)", OR, CI_lower, CI_upper),
    p_formatted = case_when(
      p.value < 0.001 ~ "<0.001",
      p.value < 0.01 ~ sprintf("%.3f", p.value),
      TRUE ~ sprintf("%.2f", p.value)
    ),
    sig_mark = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    Variable = case_when(
      grepl("SEX", term) ~ paste0("Sex: ", gsub("SEX", "", term), " vs. Female"),
      grepl("RACE", term) ~ paste0("Race: ", gsub("RACE", "", term), " vs. White"),
      grepl("CANCERTYPE", term) ~ paste0("Cancer: ", gsub("CANCERTYPE", "", term), " vs. Thoracic"),
      grepl("ECOG", term) ~ paste0("ECOG: ", ifelse(gsub("ECOG", "", term) == "2", "2/3", gsub("ECOG", "", term)), " vs. ECOG 0"),
      grepl("IV", term) ~ paste0("IV: ", gsub("IV", "", term), " vs. Chemo"),
      grepl("PO", term) ~ paste0("PO: ", gsub("PO", "", term), " vs. None or Unclassified"),
      TRUE ~ term
    )
  )

# Make Forest plot
tabletext <- cbind(
  c("Variable", result.dis$Variable),
  c("OR (95% CI)", result.dis$OR_CI),
  c("P Value", paste0(result.dis$p_formatted,result.dis$sig_mark))
)

tiff("plot/supplement2c.tiff", 
     width = 6, height = 4.3, units = "in", res = 300)

forestplot(
  labeltext = tabletext,
  mean = c(NA, result.dis$OR),
  lower = c(NA, result.dis$CI_lower),
  upper = c(NA, result.dis$CI_upper),
  is.summary = c(TRUE, rep(FALSE, nrow(result.dis))),
  align = c("l","l","c"),
  
  # X-axis format
  xlog = TRUE,
  xticks = c(0.1,1,10),
  xticks.digits = 1,
  xlab = "Odds Ratio (OR)",
  
  # Color
  col = fpColors(
    box = "black",
    line = "black" ,
    zero = "grey50"),
  
  # Line and Point format
  lwd.xaxis = 1.5,
  lwd.ci = 1.5,
  lwd.zero = 1, #OR = 1 line
  boxsize = 0.25,
  vertices = TRUE,
  
  # Font Format
  txt_gp = fpTxtGp(
    label = gpar(fontsize = 10),
    ticks = gpar(fontsize = 20),
    xlab = gpar(fontsize = 16,fontface = "bold"),
    title = gpar(
      fontsize = 13,           
      fontface = "bold",   
      lineheight = 10)),
  
  # Layout setting
  graphwidth = unit(1.5, "inches"),
  colgap = unit(2, "mm"),
  lineheight = unit(6, "mm")
)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.08, 
  gp = gpar(cex = 0.7, fontface = "italic")
)

dev.off()

```

### 3.1.6.5 Hospitalization forest plot
```{r, include=FALSE, echo=FALSE}
logis.hos <- glm(HOSPITAL ~ SEX + RACE + CANCERTYPE + ECOG + IV + PO,
                 data = demo.clean,
                 family = binomial)

summary(logis.hos)

result.hos <- tidy(logis.hos, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    OR = estimate,
    CI_lower = conf.low,
    CI_upper = conf.high,
    OR_CI = sprintf("%.2f (%.2f-%.2f)", OR, CI_lower, CI_upper),
    p_formatted = case_when(
      p.value < 0.001 ~ "<0.001",
      p.value < 0.01 ~ sprintf("%.3f", p.value),
      TRUE ~ sprintf("%.2f", p.value)
    ),
    sig_mark = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    Variable = case_when(
      grepl("SEX", term) ~ paste0("Sex: ", gsub("SEX", "", term), " vs. Female"),
      grepl("RACE", term) ~ paste0("Race: ", gsub("RACE", "", term), " vs. White"),
      grepl("CANCERTYPE", term) ~ paste0("Cancer: ", gsub("CANCERTYPE", "", term), " vs. Thoracic"),
      grepl("ECOG", term) ~ paste0("ECOG: ", ifelse(gsub("ECOG", "", term) == "2", "2/3", gsub("ECOG", "", term)), " vs. ECOG 0"),
      grepl("IV", term) ~ paste0("IV: ", gsub("IV", "", term), " vs. Chemo"),
      grepl("PO", term) ~ paste0("PO: ", gsub("PO", "", term), " vs. None or Unclassified"),
      TRUE ~ term
    )
  )

# Make Forest plot
tabletext <- cbind(
  c("Variable", result.hos$Variable),
  c("OR (95% CI)", result.hos$OR_CI),
  c("P Value", paste0(result.hos$p_formatted,result.hos$sig_mark))
)

tiff("plot/supplement2d.tiff", 
     width = 6, height = 4.3, units = "in", res = 300)

forestplot(
  labeltext = tabletext,
  mean = c(NA, result.hos$OR),
  lower = c(NA, result.hos$CI_lower),
  upper = c(NA, result.hos$CI_upper),
  is.summary = c(TRUE, rep(FALSE, nrow(result.hos))),
  align = c("l","l","c"),
  
  # X-axis format
  xlog = TRUE,
  xticks = c(0.1,1,10),
  xticks.digits = 1,
  xlab = "Odds Ratio (OR)",
  
  # Color
  col = fpColors(
    box = "black",
    line = "black" ,
    zero = "grey50"),
  
  # Line and Point format
  lwd.xaxis = 1.5,
  lwd.ci = 1.5,
  lwd.zero = 1, #OR = 1 line
  boxsize = 0.25,
  vertices = TRUE,
  
  # Font Format
  txt_gp = fpTxtGp(
    label = gpar(fontsize = 10),
    ticks = gpar(fontsize = 20),
    xlab = gpar(fontsize = 16,fontface = "bold"),
    title = gpar(
      fontsize = 13,           
      fontface = "bold",   
      lineheight = 10)),
  
  # Layout setting
  graphwidth = unit(1.5, "inches"),
  colgap = unit(2, "mm"),
  lineheight = unit(6, "mm")
)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.08, 
  gp = gpar(cex = 0.7, fontface = "italic")
)

dev.off()

```

### 3.1.6.6 Surival forest plot (Cox model)
```{r, include=FALSE, echo=FALSE}

cox_os <- coxph(
  Surv(SURVIVAL_TIME, as.numeric(as.character(DEATH))) ~ SEX + RACE + CANCERTYPE + ECOG + IV + PO ,
  data = demo.clean
)

summary(cox_os) 

cox_test <- cox.zph(cox_os)
cox_test # check assumptions for proportional hazard

result.death <- tidy(cox_os, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    HR = estimate,
    CI_lower = conf.low,
    CI_upper = conf.high,
    HR_CI = sprintf("%.2f (%.2f-%.2f)", HR, CI_lower, CI_upper),
    p_formatted = case_when(
      p.value < 0.001 ~ "<0.001",
      p.value < 0.01 ~ sprintf("%.3f", p.value),
      TRUE ~ sprintf("%.2f", p.value)
    ),
    sig_mark = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    Variable = case_when(
      grepl("SEX", term) ~ paste0("Sex: ", gsub("SEX", "", term), " vs. Female"),
      grepl("RACE", term) ~ paste0("Race: ", gsub("RACE", "", term), " vs. White"),
      grepl("CANCERTYPE", term) ~ paste0("Cancer: ", gsub("CANCERTYPE", "", term), " vs. Thoracic"),
      grepl("ECOG", term) ~ paste0("ECOG: ", ifelse(gsub("ECOG", "", term) == "2", "2/3", gsub("ECOG", "", term)), " vs. ECOG 0"),
      grepl("IV", term) ~ paste0("IV: ", gsub("IV", "", term), " vs. Chemo"),
      grepl("PO", term) ~ paste0("PO: ", gsub("PO", "", term), " vs. None or Unclassified"),
      TRUE ~ term
    )
  )

# Make Forest plot
tabletext <- cbind(
  c("Variable", result.death$Variable),
  c("HR (95% CI)", result.death$HR_CI),
  c("P Value", paste0(result.death$p_formatted,result.death$sig_mark))
)

tiff("plot/figure1a.tiff", 
     width = 6.2, height = 4.3, units = "in", res = 300)

forestplot(
  labeltext = tabletext,
  mean = c(NA, result.death$HR),
  lower = c(NA, result.death$CI_lower),
  upper = c(NA, result.death$CI_upper),
  is.summary = c(TRUE, rep(FALSE, nrow(result.death))),
  align = c("l","l","c"),
  #title = "Multivariable Cox Regression - Overall Survival",
  
  # X-axis format
  xlog = TRUE,
  xticks = c(0.1,1,10),
  xlab = "Hazard Ratio (HR)",
  
  # Color
  col = fpColors(
    box = "black",
    line = "black" ,
    zero = "grey50"),
  
  # Line and Point format
  lwd.xaxis = 1.5,
  lwd.ci = 1.5,
  lwd.zero = 1, #OR = 1 line
  boxsize = 0.25,
  vertices = TRUE,
  
  # Font Format
  txt_gp = fpTxtGp(
    label = gpar(fontsize = 11),
    ticks = gpar(fontsize = 20),
    xlab = gpar(fontsize = 16,fontface = "bold"),
    title = gpar(
      fontsize = 13,           
      fontface = "bold",   
      lineheight = 10)),
  
  # Layout setting
  graphwidth = unit(1.5, "inches"),
  colgap = unit(2, "mm"),
  lineheight = unit(6, "mm")
)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.08, 
  gp = gpar(cex = 0.7, fontface = "italic")
)

dev.off()

```

## 3.1.7 Random split into 80/20 training and testing cohorts

```{r, include=FALSE, echo=FALSE}
set.seed(123)
PTID <- unique(no.missing$PTID)
train_id <- sample(PTID, size = floor(0.8 * length(PTID)))
vali_id <- PTID[!PTID %in% train_id]

#--------------------------------------
# IRT Model
#--------------------------------------
train_data <- no.missing[no.missing$PTID %in% train_id, ]
test_data  <- no.missing[!no.missing$PTID %in% train_id, ]

train_base <- train_data %>%
            filter(TAFR == 0)

#--------------------------------------
# PRO total model
#--------------------------------------
# This model doesnt need to combine level 3 and level 4 but need further imputation on missing data
pro_beta <- read.csv(here("data","fulldata_IRT.csv")) %>%
            # re-code the binary outcome variable
            mutate(DISCONTINUE = case_when(is.na(DISCONTINUE) ~ "0",TRUE ~ "1")) %>%
            mutate(DEATH = case_when(is.na(DEATH) ~ "0",TRUE ~ "1")) %>%
        # regroup the IV and PO treatment groups same as IRT
        mutate(REGIMEN_IV = case_when(    
         grepl("Chemotherapy only",REGIMEN_IV,ignore.case = TRUE) ~ "1",
         grepl("Targeted only",REGIMEN_IV,ignore.case = TRUE) ~ "2",
         grepl("Immunotherapy only",REGIMEN_IV,ignore.case = TRUE) ~ "3",
         REGIMEN_IV %in% c("Chemo_Targeted","three combine","Chemo_Immunotherapy",
                      "Immuno_Targeted") ~ "4",
         TRUE ~"5")) %>%
  
        mutate(REGIMEN_PO = case_when(
         grepl("TKI",REGIMEN_PO,ignore.case = TRUE) ~ "1",
         TRUE ~"2")) %>%
  
        mutate(across(c(GENDER, ETHNICITY, RACE, CANCERTYPE, ECOG,REGIMEN_IV, REGIMEN_PO, C12_DOSE_REDUCTION,
                DISCONTINUE,PROCTCAE, VAR, ER, HOSPITAL,REGIMEN_IV,REGIMEN_PO),as.factor)) %>%
        select(-CONSENT_DT) %>%
        rename(SEX = GENDER, IV = REGIMEN_IV, PO = REGIMEN_PO, SURVIVAL_TIME = DEATH.DATE) %>%
        select(PTID,PROCTCAE,VAR,TIME,TAFR,DV,MDV,ER,HOSPITAL,C12_DOSE_REDUCTION,DISCONTINUE,DEATH,SURVIVAL_TIME,
               everything())

train_data_total <- pro_beta %>% filter(PTID %in% train_id)
test_data_total  <- pro_beta %>% filter(PTID %in% vali_id)

```

## 3.1.8 Export the datasets for IRT and PRO_total model (needs further imputation on missing data and data exploratory)
IRT dataset are ready for monolix code, but 

```{r, include=FALSE, echo=FALSE}
# Export the datasets for IRT and PRO total model
# IRT model - use the data frame without missing data
#write.csv(train_data, "full_IRT_training.csv", row.names = FALSE)
#write.csv(train_base, "base_IRT_training.csv", row.names = FALSE)
#write.csv(test_data, "full_IRT_testing.csv", row.names = FALSE)

# PRO total model - use the data frame with missing data (further imputation in other file)  
#write.csv(train_data_total, "pro_missing_train.csv", row.names = FALSE)
#write.csv(test_data_total, "pro_missing_test.csv", row.names = FALSE)  
```



