---
title: "Data exploratory - PRO Overll Score"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(rstatix)
library(ggpubr)
library(stringr)
library(here)
library(ggsci)

```

# 2. Import the raw data 
This analysis focus on the full data set rather than training or testing cohort only (this will be explored in another
code file for both IRT and PRO total model).

```{r, echo=FALSE}
# train
train <- read.csv(here("data","PROtotal_data","pro_missing_train.csv"))
train_id <- unique(train$PTID)
# validation
validation <- read.csv(here("data","PROtotal_data","pro_missing_test.csv"))
vali_id <- unique(validation$PTID)
```

# 3.1. Data exploratory

## 3.1.1 data imputation
last observation carry forward method
limitation of current method：
1. if a patient no response to all questions at specific time points (it is still be LOCF);
2. The filling strategy for first time was based on mode of the full trajectory of the patient.
further sensitivity analysis will be done if needed.

```{r, echo=FALSE}
#--------------------------------------
# Missing imputation with LOCF
#--------------------------------------
# combine the train and testing cohort
full_data <- rbind(train,validation) %>%
            select(-MDV) %>%
            # high missingness and variability after TAFR 360 days
            filter(TAFR <= 360)  %>%
            mutate(FIRSTFLAG = as.numeric(TIME == FIRSTTIME)) %>%
            mutate(HOSPITAL = as.factor(HOSPITAL), ER = as.factor(ER), DISCONTINUE =as.factor(DISCONTINUE),
                 C12_DOSE_REDUCTION = as.factor(C12_DOSE_REDUCTION))


# define the mode function
get_mode <- function(x) {
  x <- x[!is.na(x)]     
  if (length(x) == 0) { 
    return(NA)
  }
  uniq <- unique(x)
  uniq[which.max(tabulate(match(x, uniq)))]
}

# summarize the missing percentage pre time 
pro_missing_check <- full_data %>%
               select(PTID,PROCTCAE,TAFR,DV) %>%
               group_by(PTID, TAFR) %>%
               summarise(all_na = all(is.na(DV)),.groups = "drop") %>%
                    group_by(TAFR) %>%
                    summarise(total_patients = n(), n_all_na = sum(all_na),
                    pct_all_na = n_all_na / total_patients * 100,
                    .groups = "drop")

# impute the full original data
#some patients miss all PRO Score of item 5,6,8,13
full_fill <- full_data %>%
        group_by(PTID,PROCTCAE) %>%
        mutate(pro_mode = get_mode(DV)) %>%
        mutate(DV = ifelse(FIRSTFLAG & is.na(DV), pro_mode, DV)) %>%
        arrange(TIME) %>%
        fill(DV, .direction = "down") %>%
        ungroup() %>%
        select(-VAR,-pro_mode,-PROCTCAE)

# calculate the PRO total score and covert to 0-100 boundary data
# SCORE = total pro score; SCORE2 = transformed SCORE (0-100); SCORE3: change from baseline
full.transform <- full_fill %>%
      group_by(PTID,TIME) %>%
      summarise(
        count = sum(!is.na(DV)),
        SCORE = sum(DV,na.rm = TRUE),
        SCORE2 = SCORE/(count*4)*100,
        .groups = "drop"
      )

full_final <- full_fill %>%
          group_by(PTID,TIME) %>%
          slice(1) %>% 
          ungroup() %>%
          left_join(full.transform, by = c("PTID","TIME")) %>%
          select(-count,-DV) %>%
          group_by(PTID) %>%
          mutate(BASE = SCORE2[TAFR == 0][1]) %>%  
          ungroup() %>%
          mutate(SCORE3 = SCORE2 - BASE) %>%
          select(PTID,TIME,TAFR,SCORE,SCORE2,SCORE3,BASE,everything())

```

## 3.1.2 Total PRO score longitudinal change
```{r, echo=FALSE}
# calculate the pro mean and sd for each window and generate the plots
breaks <- c(seq(0, 360, by = 60))    
win_labs <- as.character(breaks[-1])

full_mean <- full_final %>%
    mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
    group_by(window) %>%
    summarise(
      mean = mean(SCORE2, na.rm = TRUE),
      sd = sd(SCORE2, na.rm = TRUE),
      se = sd/sqrt(n()),
      .groups = "drop")

# PRO total longitudinal change with time
ggplot(full_mean, aes(x = window, y = mean, group = 1)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),width = 0.3, size = 0.8) +
  scale_y_continuous(breaks = seq(0, 45, 15), limits = c(0, 45)) +
  labs(
    title = "Total PRO Longitudinal Change with Time",
    x = "Day",
    y = "Score (Mean ± SD)") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "lightgray"),
    legend.position = "top",
    legend.text = element_text(size = 12) 
  )

```
## 3.1.3 Total PRO score change from baseline
```{r, echo=FALSE}
full_base_mean <- full_final %>%
    mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
    group_by(window) %>%
    summarise(
      mean = mean(SCORE3, na.rm = TRUE),
      sd = sd(SCORE3, na.rm = TRUE),
      se = sd/sqrt(n()),
      .groups = "drop"
    )

# PRO total longitudinal change from baseline with time
ggplot(full_base_mean, aes(x = window, y = mean, group = 1)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.3, size = 0.8) +
  scale_y_continuous(breaks = seq(-15, 15, 10), limits = c(-15,15)) +
  labs(
    title = "Total PRO Longitudinal Change from Baseline with Time",
    x = "Day",
    y = "Score (Mean ± SD)") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "lightgray"),
    legend.position = "top",
    legend.text = element_text(size = 12) 
  )

```
## 3.1.4 Endpoint subgroup analysis
Compare PRO longitudinal change of patient with the event to patient without the event

### 3.1.4.1 Dose Discontinuation
```{r, echo=FALSE}
pro_change_dis <- full_final %>%
  mutate(DISCONTINUE = as.factor(DISCONTINUE)) %>%
  mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
  group_by(window,DISCONTINUE) %>%
  summarise(
    count = n(),
    n_pt = n_distinct(PTID),
    mean = mean(SCORE2, na.rm = TRUE),
    sd = sd(SCORE2, na.rm = TRUE),
    se = sd/sqrt(n()),
    .groups = "drop")

pro_dis <- ggplot(pro_change_dis,aes(x = window, y = mean, color = DISCONTINUE, group = DISCONTINUE)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = 0.3, size = 1) +
  #geom_text(aes(label = paste0("n=", n_pt)),
  #          vjust = -0.8, size = 3.3, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 40, 20), limits = c(0, 40)) +
  scale_color_manual(
    values = c("0" = "#1964B0", "1" = "#DB5829"),
    labels = c("1" = "Has Treatment Discontinuation",
               "0" = "No Treatment Discontinuation"),name = NULL)+
  labs(
    x = "Time (days)",
    y = "Score (Mean ± SD)"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "top",
    legend.text = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold")
  )

print(pro_dis)

ggsave("plot/supplement3b.tiff",
       plot = pro_dis,
       width = 7.5,
       height = 6,
       units = "in",
       dpi = 300)

```

### 3.1.4.2 hospitalization

```{r, echo=FALSE}
pro_change_hos <- full_final %>%
  mutate(HOSPITAL = as.factor(HOSPITAL)) %>%
  mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
  group_by(window,HOSPITAL) %>%
  summarise(
    count = n(),
    n_pt = n_distinct(PTID),
    mean = mean(SCORE2, na.rm = TRUE),
    sd = sd(SCORE2, na.rm = TRUE),
    se = sd/sqrt(n()),
    .groups = "drop")

pro_hos <- ggplot(pro_change_hos,aes(x = window, y = mean, color = HOSPITAL, group = HOSPITAL)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = 0.3, size = 1) +
  #geom_text(aes(label = paste0("n=", n_pt)),
  #          vjust = -0.8, size = 3.3, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 40, 20), limits = c(0, 40)) +
  scale_color_manual(
    values = c("0" = "#1964B0", "1" = "#DB5829"),
    labels = c("1" = "Has Hospitalization",
               "0" = "No Hospitalization"),name = NULL)+
  labs(
    x = "Time (days)",
    y = "Score (Mean ± SD)"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "top",
    legend.text = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold")
  )

print(pro_hos)

ggsave("plot/figure1c.tiff",
       plot = pro_hos,
       width = 8,
       height = 6,
       units = "in",
       dpi = 300)


```

### 3.1.4.3 ER Visit

```{r, echo=FALSE}
pro_change_er <- full_final %>%
  mutate(ER = as.factor(ER)) %>%
  mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
  group_by(window,ER) %>%
  summarise(
    count = n(),
    n_pt = n_distinct(PTID),
    mean = mean(SCORE2, na.rm = TRUE),
    sd = sd(SCORE2, na.rm = TRUE),
    se = sd/sqrt(n()),
    .groups = "drop")

pro_er <- ggplot(pro_change_er,aes(x = window, y = mean, color = ER, group = ER)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = 0.3, size = 1) +
  #geom_text(aes(label = paste0("n=", n_pt)),
  #          vjust = -0.8, size = 3.3, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 40, 20), limits = c(0, 40)) +
  scale_color_manual(
    values = c("0" = "#1964B0", "1" = "#DB5829"),
    labels = c("1" = "Has ED Visit",
               "0" = "No ED Visit"),
    name = NULL
  )+
  labs(
    x = "Time (Day)",
    y = "Score (Mean ± SD)"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "top",
    legend.text = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold")
  )

print(pro_er)

ggsave("plot/figure1d.tiff",
       plot = pro_er,
       width = 8,
       height = 6,
       units = "in",
       dpi = 300)



```

### 3.1.4.4 Dose reduction

```{r, echo=FALSE}
pro_change_red <- full_final %>%
  mutate(C12_DOSE_REDUCTION = as.factor(C12_DOSE_REDUCTION)) %>%
  mutate(window = cut(TAFR, breaks = breaks, include.lowest = TRUE, labels = win_labs)) %>%
  group_by(window,C12_DOSE_REDUCTION) %>%
  summarise(
    count = n(),
    n_pt = n_distinct(PTID),
    mean = mean(SCORE2, na.rm = TRUE),
    sd = sd(SCORE2, na.rm = TRUE),
    se = sd/sqrt(n()),
    .groups = "drop")

pro_red <- ggplot(pro_change_red,aes(x = window, y = mean, color = C12_DOSE_REDUCTION, group = C12_DOSE_REDUCTION)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                width = 0.3, size = 1) +
  #geom_text(aes(label = paste0("n=", n_pt)),
  #          vjust = -0.8, size = 3.3, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 40, 20), limits = c(0, 40)) +
  scale_color_manual(
    values = c("0" = "#1964B0", "1" = "#DB5829"),
    labels = c("1" = "Has Dose Reduction",
               "0" = "No Dose Reduction"),
    name = NULL
  )+
  labs(
    x = "Time (Day)",
    y = "Score (Mean ± SD)"
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "top",
    legend.text = element_text(size = 16),
    axis.title = element_text(size = 16, face = "bold")
  )

print(pro_red)

ggsave("plot/supplement3a.tiff",
       plot = pro_red,
       width = 7.5,
       height = 6,
       units = "in",
       dpi = 300)


```
## 3.1.5 Export the datasets for PRO_total model 
Data ready for PRO total model

```{r, include=FALSE, echo=FALSE}

train_final_data <- full_final %>%
                filter(PTID %in% train_id)

test_final_data <- full_final %>%
                filter(PTID %in% vali_id)

# Export the datasets for PRO total model
#write.csv(train_final_data, "LOCF_train.csv", row.names = FALSE)
#write.csv(test_final_data, "LOCF_test.csv", row.names = FALSE)  

```