---
title: "Data exploratory - Generate the ML datasets including covariates and individual paramters"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(dplyr)
library(tidyr)
library(here)
library(car) 
library(pROC) 
library(ResourceSelection)
library(ggplot2)
library(grid)

```

# 2. Import the raw data 

```{r, echo=FALSE}
train <- read.csv(here("data","PROtotal_data","LOCF_train.csv")) %>%
        group_by(PTID) %>%
        distinct(PTID,.keep_all = TRUE) %>%
        select(PTID,ER,HOSPITAL,DISCONTINUE,C12_DOSE_REDUCTION,DEATH,SURVIVAL_TIME,
               AGE,ETHNICITY,RACE,HEIGHT,CANCERTYPE,IV,PO,SEX,ECOG,SCORE) %>%
        rename(REDUCTION = C12_DOSE_REDUCTION) %>%
        ungroup()

validation <- read.csv(here("data","PROtotal_data","LOCF_test.csv")) %>%
        group_by(PTID) %>%
        distinct(PTID,.keep_all = TRUE) %>%
        select(PTID,SCORE,ER,HOSPITAL,DISCONTINUE,C12_DOSE_REDUCTION,DEATH,SURVIVAL_TIME,
               AGE,ETHNICITY,RACE,HEIGHT,CANCERTYPE,IV,PO,SEX,ECOG,SCORE) %>%
        rename(REDUCTION = C12_DOSE_REDUCTION) %>%
        ungroup()

# Train dataset parameters
train_prm <- read.csv(here("data","PROtotal_data","ML for CS","train.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)

# Validation 30 days dataset parameters
vali_prm_30 <- read.csv(here("data","PROtotal_data","ML for CS","vali_30.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id)%>%
          select(-tau,-gamma0,-gamma1)

# Validation 60 days dataset parameters
vali_prm_60 <- read.csv(here("data","PROtotal_data","ML for CS","vali_60.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)

# Validation 90 days dataset parameters
vali_prm_90 <- read.csv(here("data","PROtotal_data","ML for CS","vali_90.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)

# Validation 180 days dataset parameters
vali_prm_180 <- read.csv(here("data","PROtotal_data","ML for CS","vali_180.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)

# Validation fulll dataset parameters
vali_prm_full <- read.csv(here("data","PROtotal_data","ML for CS","vali_full.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id) %>%
          select(-tau,-gamma0,-gamma1)
```

# 3.1 Generate the datasets

## 3.1.1 combine demographics with individual parameters

```{r, echo=FALSE}
train_c <- train %>%
        left_join(train_prm,by = "PTID")

vali_c_30 <- validation %>%
        left_join(vali_prm_30,by = "PTID")

vali_c_60 <- validation %>%
        left_join(vali_prm_60,by = "PTID")

vali_c_90 <- validation %>%
        left_join(vali_prm_90,by = "PTID")

vali_c_180 <- validation %>%
        left_join(vali_prm_180,by = "PTID")

vali_c_full <- validation %>%
        left_join(vali_prm_full,by = "PTID")
```

## 3.1.2 Time to Event data separation

```{r, include=FALSE, echo=FALSE}
#--------------------------------------
# Training Cohort
#--------------------------------------
train_death <- train_c %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-REDUCTION)

#--------------------------------------
# 30 day validation
#--------------------------------------
vali_30_death <- vali_c_30 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-REDUCTION)

#--------------------------------------
# 60 day validation
#--------------------------------------
vali_60_death <- vali_c_60 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-REDUCTION)

#--------------------------------------
# 90 day validation data
#--------------------------------------
vali_90_death <- vali_c_90 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-REDUCTION)

#--------------------------------------
# 180 day validation
#--------------------------------------
vali_180_death <-vali_c_180 %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DISCONTINUE)

#--------------------------------------
# full validation data
#--------------------------------------
vali_full_death <-vali_c_full %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DISCONTINUE)

```

## 3.1.3 Export the datasets for PRO_total model 

```{r, include=FALSE, echo=FALSE}

write.csv(train_death,"train_OS_total.csv",row.names = FALSE)
write.csv(vali_30_death,"vali_30_OS_total.csv",row.names = FALSE)
write.csv(vali_60_death,"vali_60_OS_total.csv",row.names = FALSE)
write.csv(vali_90_death,"vali_90_OS_total.csv",row.names = FALSE)
write.csv(vali_180_death,"vali_180_OS_total.csv",row.names = FALSE)
write.csv(vali_full_death,"vali_full_OS_total.csv",row.names = FALSE)

```

# 4. Generate the heatmap of univariable analysis

## 4.1 Parameter normalization 

```{r, echo=FALSE}
# normalized the parameters based on distribution
vars <- c("BASEP", "SLP", "Emax","Kd")
vars_base <- c("BASEP", "SLP", "Emax") #normal distribution
vars_log <- c("Kd") #lognormal distribution

train_ml <- train_c

train_ml[vars_base] <- scale(train_ml[vars_base])
train_ml[vars_log] <- scale(log(train_ml[vars_log]))

```


## 4.2 Univaraible logistic regression

### 4.2.1 Dose Discontinuation

```{r, echo=FALSE}
train_logis_dis <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("DISCONTINUE ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_dis <- rbind(train_logis_dis,
                   data.frame(
                     variable = prm,
                     P_value = coef(summary(m))[2,4],
                     OR = exp(coef(m))[2],
                     OR_low = exp(confint(m))[2,1],
                     OR_high = exp(confint(m))[2,2]
                   ))}
train_logis_dis

```

### 4.2.2 ER Visit
```{r, echo=FALSE}
train_logis_er <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("ER ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_er <- rbind(train_logis_er,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_er

```

### 4.2.3 Hospitalization
```{r, echo=FALSE}
train_logis_hos <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("HOSPITAL ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_hos <- rbind(train_logis_hos,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_hos

```

### 4.2.4 Dose Reduction
```{r, echo=FALSE}
train_logis_red <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("REDUCTION ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_red <- rbind(train_logis_red,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_red

```

### 4.2.5 Survival
```{r, echo=FALSE}
train_logis_death <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("DEATH ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_death <- rbind(train_logis_death,
                   data.frame(
                     variable = prm,
                     P_value = coef(summary(m))[2,4],
                     OR = exp(coef(m))[2],
                     OR_low = exp(confint(m))[2,1],
                     OR_high = exp(confint(m))[2,2]
                   ))}
train_logis_death
```

### 4.2.6 Heatmap
```{r, echo=FALSE}
train_logis_er$endpoint <- "ED Visit"
train_logis_dis$endpoint <- "Discontinuation"
train_logis_hos$endpoint <- "Hopitalization"
train_logis_red$endpoint <- "Reduction"
train_logis_death$endpoint <- "Survival"

train_heatmap <- rbind(train_logis_er, train_logis_dis, train_logis_hos, train_logis_red,train_logis_death)

train_heatmap$logOR <- log(train_heatmap$OR)

train_heatmap <- train_heatmap %>%
  mutate(signif = case_when(
    P_value < 0.001 ~ "***",
    P_value < 0.01  ~ "**",
    P_value < 0.05  ~ "*",
    TRUE            ~ ""  
  ))

p1 <- ggplot(train_heatmap, aes(x = variable, y = endpoint, fill = log(OR))) +
  geom_tile(color = "white") +
  geom_text(aes(label = signif), color = "black", size = 8) +
  scale_fill_gradient2(
    low = "darkblue",
    mid = "white",
    high = "red",
    midpoint = 0,
    
    limits = c(-0.5, 1),   
    breaks = c(-0.5, 0, 0.5, 1),  
    labels = c("-0.5", "0", "0.5", "1.0"),
    
    name = "Log(OR)",
    
    guide = guide_colorbar(
      title.position = "top",
      barwidth = 1,              
      barheight = 8,            
      frame.colour = "black",    
      ticks.colour = "black"     
    )
  ) +
  
  coord_equal() +
  labs(x = NULL, y = NULL) +
  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15, color = "black"),
    axis.text.y = element_text(size = 15.5, color = "black"),
    axis.title = element_text(size = 15.5),
    
    legend.title = element_text(size = 15, face = "bold"),  
    legend.text = element_text(size = 15),                  
    legend.position = "right",
    
    plot.margin = margin(5, 5, 5, 5, "mm")
  )

print(p1)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.15, 
  gp = gpar(cex = 0.8, fontface = "italic")
)
```


