---
title: "Data exploratory - Generate the ML datasets including covariates and individual paramters"
author: "Congyu Zhang"
output: html_document
---
# 1. Load packages and set up environment

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())

library(dplyr)
library(tidyr)
library(here)
library(car) 
library(pROC) 
library(ResourceSelection)
library(ggplot2)
library(grid)

```

# 2. Import the raw data 

```{r, echo=FALSE}
train <- read.csv(here("data","IRT_data","ML_data","full_IRT_training.csv")) %>%
        group_by(PTID) %>%
        distinct(PTID,.keep_all = TRUE) %>%
        select(PTID,ER,HOSPITAL,DISCONTINUE,C12_DOSE_REDUCTION,DEATH,SURVIVAL_TIME,
               AGE,ETHNICITY,RACE,HEIGHT,CANCERTYPE,IV,PO,SEX,ECOG) %>%
        rename(REDUCTION = C12_DOSE_REDUCTION) %>%
        ungroup()

validation <- read.csv(here("data","IRT_data","ML_data","full_IRT_testing.csv")) %>%
        group_by(PTID) %>%
        distinct(PTID,.keep_all = TRUE) %>%
        select(PTID,ER,HOSPITAL,DISCONTINUE,C12_DOSE_REDUCTION,DEATH,SURVIVAL_TIME,
               AGE,ETHNICITY,RACE,HEIGHT,CANCERTYPE,IV,PO,SEX,ECOG) %>%
        rename(REDUCTION = C12_DOSE_REDUCTION) %>%
        ungroup()

# Train dataset parameters
train_prm <- read.csv(here("data","IRT_data","ML_data",
                           "model_output","train.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation 30 days dataset parameters
vali_prm_30 <- read.csv(here("data","IRT_data","ML_data",
                             "model_output3","vali30.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation 60 days dataset parameters
vali_prm_60 <- read.csv(here("data","IRT_data","ML_data","model_output3","vali60.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation 90 days dataset parameters
vali_prm_90 <- read.csv(here("data","IRT_data","ML_data","model_output3","vali90.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation 180 days dataset parameters
vali_prm_180 <- read.csv(here("data","IRT_data","ML_data","model_output3","vali180.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation full dataset parameters
vali_prm_full <- read.csv(here("data","IRT_data","ML_data","model_output3","full.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation hospitalization dataset parameters
vali_prm_hos <- read.csv(here("data","IRT_data","ML_data","model_output3","hospital.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation ER dataset parameters
vali_prm_er <- read.csv(here("data","IRT_data","ML_data","model_output3","er.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 

# Validation Reduction dataset parameters
vali_prm_red <- read.csv(here("data","IRT_data","ML_data","model_output3","reduction.txt")) %>% 
          select(id,contains("SAEM")) %>% 
          select(-matches("DIS([1-9]|1[0-4])")) %>%
          select(-matches("Q([1-9]|1[0-4])th1")) %>%
          select(-matches("Q([1-9]|1[0-4])th2")) %>%
          select(-matches("Q([1-9]|1[0-4])th3")) %>%
          rename_with(~gsub("_SAEM$", "", .), contains("SAEM")) %>%
          rename(PTID = id,EMAXGI = PROGGI,EMAXPS = PROGPS,EMAXPD = PROGPD) 
```

# 3.1 Generate the datasets

## 3.1.1 combine demographics with individual parameters

```{r, echo=FALSE}
train_c <- train %>%
        left_join(train_prm,by = "PTID")

vali_c_30 <- validation %>%
        left_join(vali_prm_30,by = "PTID")

vali_c_60 <- validation %>%
        left_join(vali_prm_60,by = "PTID")

vali_c_90 <- validation %>%
        left_join(vali_prm_90,by = "PTID")

vali_c_180 <- validation %>%
        left_join(vali_prm_180,by = "PTID")

vali_c_full <- validation %>%
        left_join(vali_prm_full,by = "PTID")

vali_c_hos <- validation %>%
        left_join(vali_prm_hos,by = "PTID")

vali_c_er <- validation %>%
        left_join(vali_prm_er,by = "PTID")

vali_c_red <- validation %>%
        left_join(vali_prm_red,by = "PTID")
```

## 3.1.2 Time to Event data separation
### 3.1.2.1 Training Cohort
```{r, include=FALSE, echo=FALSE}
train_er <- train_c %>%
          select(-DISCONTINUE,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
train_hos <- train_c %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-DEATH,-SURVIVAL_TIME)
  
train_dis <-train_c %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
train_red <- train_c %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

train_death <- train_c %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-REDUCTION)
```

### 3.1.2.2 Testing 30 days
```{r, include=FALSE, echo=FALSE}
vali_30_er <- vali_c_30 %>%
          select(-DISCONTINUE,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
vali_30_hos <- vali_c_30 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-DEATH,-SURVIVAL_TIME)

vali_30_dis <- vali_c_30 %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
vali_30_red <- vali_c_30 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_30_death <- vali_c_30 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-HOSPITAL)
```

### 3.1.2.3 Testing 60 days
```{r, include=FALSE, echo=FALSE}
vali_60_er <- vali_c_60 %>%
          select(-DISCONTINUE,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
vali_60_hos <- vali_c_60 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-DEATH,-SURVIVAL_TIME)
  
vali_60_dis <-vali_c_60 %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
vali_60_red <- vali_c_60 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_60_death <- vali_c_60 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-HOSPITAL)
```

### 3.1.2.4 Testing 90 days
```{r, include=FALSE, echo=FALSE}
vali_90_er <- vali_c_90 %>%
          select(-DISCONTINUE,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
  
vali_90_hos <- vali_c_90 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-DEATH,-SURVIVAL_TIME)
  
vali_90_dis <-vali_c_90 %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_90_red <- vali_c_90 %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_90_death <- vali_c_90 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-HOSPITAL)
```

### 3.1.2.5 Testing 180 days
```{r, include=FALSE, echo=FALSE}
vali_180_dis <-vali_c_180 %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_180_death <- vali_c_180 %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-HOSPITAL)
```

### 3.1.2.6 Testing Full trajectory
```{r, include=FALSE, echo=FALSE}
vali_full_dis <-vali_c_full %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

vali_full_death <-vali_c_full %>%
          select(-ER,-REDUCTION,-HOSPITAL,-DISCONTINUE)
```

### 3.1.2.7 Testing before event
```{r, include=FALSE, echo=FALSE}
#---hospitalization data set
vali_hospital <- vali_c_hos %>%
          select(-DISCONTINUE,-REDUCTION,-ER,-DEATH,-SURVIVAL_TIME)

#---ER data set
vali_er <- vali_c_er %>%
          select(-DISCONTINUE,-REDUCTION,-HOSPITAL,-DEATH,-SURVIVAL_TIME)

#---Dose reduction data set
vali_reduction <- vali_c_red %>%
          select(-DISCONTINUE,-ER,-HOSPITAL,-DEATH,-SURVIVAL_TIME)
```

## 3.1.3 Export the datasets for PRO_total model 

```{r, include=FALSE, echo=FALSE}
write.csv(train_er,"train_er.csv",row.names = FALSE)
write.csv(train_dis,"train_dis.csv",row.names = FALSE)
write.csv(train_hos,"train_hos.csv",row.names = FALSE)
write.csv(train_red,"train_red.csv",row.names = FALSE)
write.csv(train_death,"train_OS.csv",row.names = FALSE)

write.csv(vali_30_er,"vali_30_er.csv",row.names = FALSE)
write.csv(vali_30_hos,"vali_30_hos.csv",row.names = FALSE)
write.csv(vali_30_red,"vali_30_red.csv",row.names = FALSE)
write.csv(vali_30_dis,"vali_30_dis.csv",row.names = FALSE)
write.csv(vali_30_death,"vali_30_OS.csv",row.names = FALSE)

write.csv(vali_60_er,"vali_60_er.csv",row.names = FALSE)
write.csv(vali_60_hos,"vali_60_hos.csv",row.names = FALSE)
write.csv(vali_60_dis,"vali_60_dis.csv",row.names = FALSE)
write.csv(vali_60_red,"vali_60_red.csv",row.names = FALSE)
write.csv(vali_60_death,"vali_60_OS.csv",row.names = FALSE)

write.csv(vali_90_er,"vali_90_er.csv",row.names = FALSE)
write.csv(vali_90_hos,"vali_90_hos.csv",row.names = FALSE)
write.csv(vali_90_dis,"vali_90_dis.csv",row.names = FALSE)
write.csv(vali_90_red,"vali_90_red.csv",row.names = FALSE)
write.csv(vali_90_death,"vali_90_OS.csv",row.names = FALSE)

write.csv(vali_180_death,"vali_180_OS.csv",row.names = FALSE)
write.csv(vali_180_dis,"vali_180_dis.csv",row.names = FALSE)

write.csv(vali_full_death,"vali_full_OS.csv",row.names = FALSE)
write.csv(vali_full_dis,"vali_full_dis.csv",row.names = FALSE)

write.csv(vali_hospital,"vali_hos_event.csv",row.names = FALSE)
write.csv(vali_er,"vali_er_event.csv",row.names = FALSE)
write.csv(vali_reduction,"vali_red_event.csv",row.names = FALSE)
```

# 4. Generate the heatmap of univariable analysis

## 4.1 Parameter normalization 

```{r, echo=FALSE}
# normalized the parameters based on distribution
vars <- c("BASEGI","EMAXGI","KDGI","SLPGI",
          "BASEPS","EMAXPS","KDPS","SLPPS",
          "BASEPD","EMAXPD","KDPD","SLPPD")

vars_base <- c("BASEGI", "BASEPS", "BASEPD")    #normal distribution
vars_log <- c("EMAXGI","KDGI","SLPGI",
              "EMAXPS","KDPS","SLPPS",
              "EMAXPD","KDPD","SLPPD")      #lognormal distribution

#normal distribution
vars_base <- c("BASEGI", "BASEPS", "BASEPD",
               "EMAXGI","SLPGI","EMAXPS","SLPPS","EMAXPD","SLPPD")

vars_log <- c("KDGI","KDPS","KDPD")      #lognormal distribution


train_ml <- train_c

train_ml[vars_base] <- scale(train_ml[vars_base])
train_ml[vars_log] <- scale(log(train_ml[vars_log]))

```


## 4.2 Univaraible logistic regression

### 4.2.1 Dose Discontinuation

```{r, echo=FALSE}
# Univaraite logistic regression
train_logis_dis <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("DISCONTINUE ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_dis <- rbind(train_logis_dis,
                   data.frame(
                     variable = prm,
                     P_value = coef(summary(m))[2,4],
                     OR = exp(coef(m))[2],
                     OR_low = exp(confint(m))[2,1],
                     OR_high = exp(confint(m))[2,2]
                   ))}
train_logis_dis

# multivariable logistic regression
dis_formula <- as.formula(
  paste("DISCONTINUE ~", paste(vars, collapse = " + "))
)

dis_multi <- glm(dis_formula, data = train_ml, family = binomial)
summary_dis <- summary(dis_multi)

train_logis_dis2 <- data.frame(
  variable = rownames(coef(summary_dis))[-1],  
  P_value = coef(summary_dis)[-1, 4],
  OR = exp(coef(dis_multi))[-1],
  OR_low = exp(confint(dis_multi))[-1, 1],
  OR_high = exp(confint(dis_multi))[-1, 2]
)

```

### 4.2.2 ER Visit
```{r, echo=FALSE}
train_logis_er <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("ER ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_er <- rbind(train_logis_er,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_er

```

### 4.2.3 Hospitalization
```{r, echo=FALSE}
train_logis_hos <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("HOSPITAL ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_hos <- rbind(train_logis_hos,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_hos

```

### 4.2.4 Dose Reduction
```{r, echo=FALSE}
train_logis_red <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("REDUCTION ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_red <- rbind(train_logis_red,
                           data.frame(
                             variable = prm,
                             P_value = coef(summary(m))[2,4],
                             OR = exp(coef(m))[2],
                             OR_low = exp(confint(m))[2,1],
                             OR_high = exp(confint(m))[2,2]
                           ))}
train_logis_red

```

### 4.2.5 Survival
```{r, echo=FALSE}
train_logis_death <- data.frame()
# check single variable p-value check
for(prm in vars){
  f <- as.formula(paste("DEATH ~", prm))
  m <- glm(f, data=train_ml, family=binomial)
  train_logis_death <- rbind(train_logis_death,
                   data.frame(
                     variable = prm,
                     P_value = coef(summary(m))[2,4],
                     OR = exp(coef(m))[2],
                     OR_low = exp(confint(m))[2,1],
                     OR_high = exp(confint(m))[2,2]
                   ))}
train_logis_death
```

### 4.2.6 Heatmap
```{r, echo=FALSE}
train_logis_er$endpoint <- "ED Visit"
train_logis_dis$endpoint <- "Discontinuation"
train_logis_hos$endpoint <- "Hopitalization"
train_logis_red$endpoint <- "Reduction"
train_logis_death$endpoint <- "Survival"

train_heatmap <- rbind(train_logis_er, train_logis_dis, train_logis_hos, train_logis_red,train_logis_death)

train_heatmap$logOR <- log(train_heatmap$OR)

train_heatmap <- train_heatmap %>%
  mutate(signif = case_when(
    P_value < 0.001 ~ "***",
    P_value < 0.01  ~ "**",
    P_value < 0.05  ~ "*",
    TRUE            ~ ""  
  ))

p1 <- ggplot(train_heatmap, aes(x = variable, y = endpoint, fill = log(OR))) +
  geom_tile(color = "white") +
  geom_text(aes(label = signif), color = "black", size = 8) +
  scale_fill_gradient2(
    low = "darkblue",
    mid = "white",
    high = "red",
    midpoint = 0,
    
    limits = c(-0.5, 1.2),   
    breaks = c(-0.5, 0, 0.5, 1),  
    labels = c("-0.5", "0", "0.5", "1.0"),
    
    name = "Log(OR)",
    
    guide = guide_colorbar(
      title.position = "top",
      barwidth = 1,              
      barheight = 8,            
      frame.colour = "black",    
      ticks.colour = "black"     
    )
  ) +
  
  coord_equal() +
  labs(x = NULL, y = NULL) +
  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15, color = "black"),
    axis.text.y = element_text(size = 15.5, color = "black"),
    axis.title = element_text(size = 15.5),
    
    legend.title = element_text(size = 15, face = "bold"),  
    legend.text = element_text(size = 15),                  
    legend.position = "right",
    
    plot.margin = margin(5, 5, 5, 5, "mm")
  )

print(p1)

grid::grid.text(
  "* P < 0.05, ** P < 0.01, *** P < 0.001", 
  x = 0.2, y = 0.15, 
  gp = gpar(cex = 0.8, fontface = "italic")
)

ggsave("plot/figure4a_normal.tiff",
       plot = p1,
       width = 14,
       height = 6,
       units = "in",
       dpi = 300)

```


